---
title: "Building and Validating Predictive Models for Acute Kidney Injury (AKI) using PCORnet CDM (Part I)"
author: "xing song"
date: "August 23, 2018"
output: html_document
params:
  DBMS_type: Oracle
  remote_CDM: FALSE
  incl_NDC: FALSE
---
### Aim 1: Data Feasibility Evaluation over GPC sites.
#### Objective 1.1: Study cohort extraction and characterization

##### Inclusion criteria:

* (IP,IS,EI) visits with length of stay (LOS) >= 2 days; 
* age at visit >= 18 years old

***

##### Exclusion criteria:

* had no documented SCr measurements during admission; 
* developed stage 2 AKI (eGFR <= 15 mL/min per 1.73m^2) initially; 
* pre-existing chronic renal failure (DX); 
* received RRT within 48 hours since admission; 
* burn patients (DRG)

***

##### AKI Staging

AKI Stages are defined based on [KDIGO](http://www.kdigo.org/clinical_practice_guidelines/pdf/KDIGO%20AKI%20Guideline.pdf):

* AKI 1: increase in serum creatinine by >=**0.3 mg/dL** within **48 hours** OR **1.5-1.9 times** baseline within **7 days**; 
* AKI 2: **2.0 to 2.9 times** baseline within **7 days**; 
* AKI 3: increase in serum creatinine to >= **4.0 mg/dL** within **48 hours** OR **3.0 times** baseline within **7 days**

***

##### Implementation: Extracting AKI Study Cohort
```{r setup, include=F}
#source utility functions
source("./R/util.R")
source("./R/extract_cohort.R")
source("./R/viz.R")

#load libraries
require_libraries(c("DBI",
                    "tidyr",
                    "dplyr",
                    "magrittr",
                    "stringr",
                    "knitr",
                    "kableExtra",
                    "ggplot2",
                    "ggrepel",
                    "RCurl",
                    "XML",
                    "openxlsx"))

```


```{r connect, include=F}
#establish the connection between r-studio and CDM server (Oracle)
config_file_path<-"./config.csv"
config_file<-read.csv(config_file_path,stringsAsFactors = F)
conn<-connect_to_db(params$DBMS_type,config_file)
```


```{r cohort,message=F}
#extract cohort --Table1
# by default, we assume cdm schema is on the same server as current schema,
cohort<-extract_cohort(conn,
                       remote_CDM=params$remote_CDM,
                       cdm_db_name=config_file$cdm_db_name,
                       cdm_db_schema=config_file$cdm_db_schema,
                       start_date="2010-01-01",
                       end_date="2018-12-31",
                       verb=F)
```

The above codes extracted AKI study cohort based on the "Inclusion" and "Exclusion" criteria specified above. The final output will be automatically saved in the current working directory `r getwd()` as "Table1.rda". More details are included in the following consort diagram.  

```{r tbl1, include=F}
Table1<-cohort$aki_enc
consort_tbl<-cohort$attrition

# collect summaries
enc_tot<-length(unique(Table1$ENCOUNTERID))

tbl1_dsa<-Table1 %>% 
  dplyr::select(PATID, ENCOUNTERID,
                NONAKI_SINCE_ADMIT, 
                AKI1_SINCE_ADMIT,
                AKI2_SINCE_ADMIT,
                AKI3_SINCE_ADMIT) %>%
  gather(stage, days_since_admit,-PATID,-ENCOUNTERID) %>%
  mutate(stage=gsub("_.*","",stage)) %>% 
  filter(!is.na(days_since_admit)) 

tbl1_summ<-tbl1_dsa %>%
  group_by(stage) %>%
  dplyr::summarize(pat_cnt=length(unique(PATID)),
                   enc_cnt=length(unique(ENCOUNTERID)),
                   min_time=min(days_since_admit,na.rm=T),
                   q1_time=quantile(days_since_admit,probs=0.25,na.rm=T),
                   median_time=median(days_since_admit,na.rm=T),
                   mean_time=round(mean(days_since_admit,na.rm=T),1),
                   q3_time=quantile(days_since_admit,probs=0.75,na.rm=T),
                   max_time=max(days_since_admit,na.rm=T),
                   sd_time=round(sd(days_since_admit,na.rm=T),2)) %>%
  mutate(semi_IQR_time=0.5*(q3_time-q1_time)) %>%
  #HIPPA, low counts masking
  mutate(pat_cnt=ifelse(as.numeric(pat_cnt)<11,"<11",as.character(pat_cnt)),
         enc_cnt=ifelse(as.numeric(enc_cnt)<11,"<11",as.character(enc_cnt)))

# tbl1_summ2<-tbl1_dsa %>%
#   mutate(dsa_bin=case_when(days_since_admit <10 ~ paste0("0",days_since_admit," days"),
#                            days_since_admit >=10 & days_since_admit < 31 ~ paste(days_since_admit,"days"),
#                            days_since_admit >=31 ~ '31 days(1mth) <')) %>%
#   group_by(stage,dsa_bin) %>%
#   dplyr::summarize(enc_cnt=length(unique(ENCOUNTERID))) %>%
#   spread(stage,enc_cnt,fill=0) %>%
#   mutate(AKI1_cum=cumsum(AKI1),
#          AKI2_cum=cumsum(AKI2),
#          AKI3_cum=cumsum(AKI3),
#          NONAKI_cum=cumsum(NONAKI)) %>%
#   arrange(desc(dsa_bin)) %>%
#   mutate(NONAKI=cumsum(NONAKI)) %>%
#   arrange(dsa_bin)

#save results
saveRDS(Table1,file="./data/Table1.rda")
saveRDS(consort_tbl,file="./data/consort_tbl.rda")

#clean up
rm(cohort); gc()
```


```{r consort, echo=F}
#consort diagram
consort_diag(consort_tbl)
```


```{r temp, include=F}
# auxilliary summaries and tables
enc_tot<-length(unique(Table1$ENCOUNTERID))
# critical dates of AKI encounters
aki_stage_ind<-Table1 %>%
  dplyr::select(PATID, ENCOUNTERID, ADMIT_DATE, DISCHARGE_DATE,
                NONAKI_ANCHOR, AKI1_ONSET,AKI2_ONSET,AKI3_ONSET) %>%
  gather(chk_pt, critical_date,-PATID,-ENCOUNTERID) %>%
  filter(!is.na(critical_date)) %>%
  mutate(chk_pt=gsub("_.*","",chk_pt)) %>%
  group_by(chk_pt) %>%
  dplyr::mutate(stg_tot_cnt=n()) %>%
  ungroup %>%
  arrange(PATID, ENCOUNTERID, chk_pt, critical_date, stg_tot_cnt)
```

***

#### Objective 1.2: Variable Collection and Summaries (Table by Table)
In this section, we will collect variables from PCORNET_CDM tables: *DEMOGRAPHIC*, *ENCOUNTER*, *VITAL*, *LAB_RESULT_CM*, *PRESCRIBING MEDICATION*, *DIAGNOSIS*, *PROCEDURE*, check data quality and generate variable summaries table by table.

##### Demographic 
Demographic variables include *Age (calculated from Birth_Date)*, *Sex*, *Race*, *Ethnicity*.  

```{r demo, include=F}
## demographic
conn<-connect_to_db(params$DBMS_type,config_file)

sql<-parse_sql(paste0("./inst/",params$DBMS_type,"/collect_demo.sql"),
                           cdm_db_link=config_file$cdm_db_link,
                           cdm_db_name=config_file$cdm_db_name,
                           cdm_db_schema=config_file$cdm_db_schema)

demo<-execute_single_sql(conn,
                         statement=sql$statement,
                         write=(sql$action=="write")) %>%
  mutate(AGE_GRP=case_when(AGE<= 25 ~ "18-25",
                           AGE >= 26 & AGE <= 35 ~ "26-35",
                           AGE >= 36 & AGE <= 45 ~ "36-45",
                           AGE >= 46 & AGE <= 55 ~ "46-55",
                           AGE >= 56 & AGE <= 65 ~ "56-65",
                           AGE >= 66 ~ "66<=")) %>%
  dplyr::select(PATID,ENCOUNTERID,
                AGE,AGE_GRP,SEX,RACE,HISPANIC,DDAYS_SINCE_ENC) %>%
  replace_na(list(AGE="NI",
                  AGE_GRP="NI",
                  SEX="NI",
                  RACE="NI",
                  HISPANIC="NI")) %>%
  gather(key,value,-PATID,-ENCOUNTERID) %>%
  unique

#collect summaries
demo_summ<-aki_stage_ind %>% 
  filter(!chk_pt %in% c("DISCHARGE")) %>%
  dplyr::select(-critical_date) %>%
  left_join(demo %>% 
              filter(!(key %in% c("AGE","DDAYS_SINCE_ENC"))), 
            by="ENCOUNTERID") %>%
  group_by(chk_pt,stg_tot_cnt,key,value) %>%
  dplyr::summarize(enc_cnt = n(),
                   enc_prop = round(n()/stg_tot_cnt[1],2)) %>%
  ungroup %>% dplyr::select(-stg_tot_cnt) %>%
  gather(summ,summ_val,-chk_pt,-key,-value) %>%
  # # decode demo_val
  # left_join(meta %>% dplyr::select(COLUMN_NAME,VAR_CODE,VAR_NAME),
  #           by=c("demo_type"="COLUMN_NAME","demo_val"="VAR_CODE")) %>%
  # dplyr::mutate(demo_val = ifelse(!is.na(VAR_NAME),VAR_NAME,demo_val)) %>%
  # dplyr::select(-VAR_NAME) %>%
  # unite("demo_type_cat",c("demo_type","demo_val")) %>%
  # attach totals at bottom
  bind_rows(aki_stage_ind %>%
              filter(!chk_pt %in% c("DISCHARGE")) %>%
              dplyr::select(chk_pt,stg_tot_cnt) %>% 
              unique %>% 
              dplyr::rename(enc_cnt=stg_tot_cnt) %>%
              mutate(enc_prop=round(enc_cnt/enc_tot,2),
                     key="TOTAL",
                     value="(%/overall)") %>%
              gather(summ,summ_val,-chk_pt,-key,-value) %>%
              dplyr::select(key,value,chk_pt,summ,summ_val)) %>%
  unite("stg_summ",c("chk_pt","summ")) %>%
  unique %>% spread(stg_summ,summ_val) %>%
  replace(.,is.na(.),0)

#save results
saveRDS(demo,file=paste0("./data/",toupper(sql$tbl_out),".rda"))
# saveRDS(demo_summ,file="./data/demo_summ.rda")

#clean up
rm(demo); gc()
```


```{r demo_out, echo=F}
demo_nice_tbl<-demo_summ %>%
  gather(summ,summ_val,-key,-value) %>%
  mutate(summ_val=ifelse(grepl("_prop",summ),paste0(summ_val*100,"%"),as.character(summ_val))) %>%
  spread(summ,summ_val) %>%
  unite("ADMIT",paste0("ADMIT_",c("enc_cnt","enc_prop")),sep=", ") %>%
  unite("AKI1",paste0("AKI1_",c("enc_cnt","enc_prop")),sep=", ") %>%
  unite("AKI2",paste0("AKI2_",c("enc_cnt","enc_prop")),sep=", ") %>%
  unite("AKI3",paste0("AKI3_",c("enc_cnt","enc_prop")),sep=", ") %>%
  unite("NONAKI",paste0("NONAKI_",c("enc_cnt","enc_prop")),sep=", ") %>%
  arrange(key,value)

kable(demo_nice_tbl,
      caption="Table1 - Demographic Summaries at AKI1, AKI2, AKI3") %>%
  kable_styling("striped", full_width = F) %>%
  group_rows("Age Group", 1,6) %>%
  group_rows("Hispanic", 7, 9) %>%
  group_rows("Race", 10, 18) %>%
  group_rows("Sex", 19, 21) %>%  
  group_rows("Total", 22, 22)  
  
```

Demographic characterizations for patients at different AKI stages are summarized in Table 1.

***


##### Vital 
Vital variables include: *Height*, *Weight*, *BMI*, *Blood Pressure (Systolic, Diastolic)*, *Smoking Status*. 

```{r vital, include=F}
## vital
conn<-connect_to_db(params$DBMS_type,config_file)

sql<-parse_sql(paste0("./inst/",params$DBMS_type,"/collect_vital.sql"),
               cdm_db_link=config_file$cdm_db_link,
               cdm_db_name=config_file$cdm_db_name,
               cdm_db_schema=config_file$cdm_db_schema)

vital<-execute_single_sql(conn,
                          statement=sql$statement,
                          write=(sql$action=="write")) %>%
  mutate(BMI_GRP = case_when(ORIGINAL_BMI <= 25 ~ "BMI <= 25",
                             ORIGINAL_BMI > 25 &  ORIGINAL_BMI <= 30 ~ "BMI 26-30",
                             ORIGINAL_BMI >=31  ~ "BMI >= 31")) %>%
  dplyr::rename(dsa=DAYS_SINCE_ADMIT,
                timestamp=MEASURE_DATE_TIME) %>%
  gather(key,value,-PATID,-ENCOUNTERID,-dsa,-timestamp) %>%
  filter(!is.na(key) & !is.na(value)) %>%
  mutate(key=recode(key,
                    ORIGINAL_BMI="BMI",
                    SYSTOLIC="BP_SYSTOLIC",
                    DIASTOLIC="BP_DIASTOLIC")) %>%
  unique

vital1<-vital %>%
  dplyr::select(ENCOUNTERID, key, value, dsa) %>%
  filter(key %in% c("HT","WT","BMI","BP_DIASTOLIC","BP_SYSTOLIC")) %>%
  mutate(value=as.numeric(value)) %>%
  mutate(param_low=case_when(key=="HT" ~ 0,
                             key=="WT" ~ 0,
                             key=="BMI" ~ 0,
                             key %in% c("BP_DIASTOLIC",
                                        "BP_SYSTOLIC") ~ 40),
         param_high=case_when(key=="HT" ~ 94.99,
                              key=="WT" ~ 350,
                              key=="BMI" ~ 50,
                              key=="BP_DIASTOLIC"~120,
                              key=="BP_SYSTOLIC" ~ 210)) %>%
  mutate(dsa_grp=case_when(dsa < 0 ~ "0>",
                           dsa >=0 & dsa < 1 ~ "1",
                           dsa >=1 & dsa < 2 ~ "2",
                           dsa >=2 & dsa < 3 ~ "3",
                           dsa >=3 & dsa < 4 ~ "4",
                           dsa >=4 & dsa < 5 ~ "5",
                           dsa >=5 & dsa < 6 ~ "6",
                           dsa >=6 & dsa < 7 ~ "7",
                           dsa >=7 ~ "7<"))


# collect summaries
vital_summ<-vital1 %>%
  group_by(key) %>%
  dplyr::summarize(record_cnt=n(),
                   enc_cnt=length(unique(ENCOUNTERID)),
                   low_cnt=sum((value<param_low)),
                   high_cnt=sum((value>param_high)),
                   min=min(value,na.rm=T),
                   mean=round(mean(value,na.rm=T)),
                   sd=round(sd(value,na.rm=T)),
                   median=round(median(value,na.rm=T)),
                   max=max(value,na.rm=T)) %>%
  ungroup %>%
  mutate(cov=round(sd/mean,1)) %>%
  #HIPPA, low counts masking
  mutate(enc_cnt=ifelse(as.numeric(enc_cnt)<11 & as.numeric(enc_cnt)>0,"<11",enc_cnt),
         record_cnt=ifelse(as.numeric(record_cnt)<11 & as.numeric(record_cnt)>0,"<11",record_cnt),
         low_cnt=ifelse(as.numeric(low_cnt)<11 & as.numeric(low_cnt)>0,"<11",as.character(low_cnt)),
         high_cnt=ifelse(as.numeric(high_cnt)<11 & as.numeric(high_cnt)>0,"<11",as.character(high_cnt))) %>%
  gather(summ,overall,-key) %>%
  mutate(summ=recode(summ,
                     enc_cnt="1.encounters#",
                     record_cnt="2.records#",
                     low_cnt="3.low_records#",
                     high_cnt="4.high_records#",
                     min="5a.min",
                     median="5b.median",
                     mean="5c.mean",
                     sd="5d.sd",
                     cov="5e.cov",
                     max="5f.max")) %>%
  left_join(
    vital1 %>%
      group_by(key,dsa_grp) %>%
      dplyr::summarize(record_cnt=n(),
                       enc_cnt=length(unique(ENCOUNTERID)),
                       low_cnt=sum((value<param_low)),
                       high_cnt=sum((value>param_high)),
                       min=min(value,na.rm=T),
                       mean=round(mean(value,na.rm=T)),
                       sd=round(sd(value,na.rm=T)),
                       median=round(median(value,na.rm=T)),
                       max=max(value,na.rm=T)) %>%
      ungroup %>%
      mutate(cov=round(sd/mean,1)) %>%
      #HIPPA, low counts masking
      mutate(enc_cnt=ifelse(as.numeric(enc_cnt)<11 & as.numeric(enc_cnt)>0,"<11",enc_cnt),
             record_cnt=ifelse(as.numeric(record_cnt)<11 & as.numeric(record_cnt)>0,"<11",record_cnt),
             low_cnt=ifelse(as.numeric(low_cnt)<11 & as.numeric(low_cnt)>0,"<11",as.character(low_cnt)),
             high_cnt=ifelse(as.numeric(high_cnt)<11 & as.numeric(high_cnt)>0,"<11",as.character(high_cnt))) %>%
      gather(summ,summ_val,-key,-dsa_grp) %>%
      spread(dsa_grp,summ_val) %>%
      mutate(summ=recode(summ,
                         enc_cnt="1.encounters#",
                         record_cnt="2.records#",
                         low_cnt="3.low_records#",
                         high_cnt="4.high_records#",
                         min="5a.min",
                         median="5b.median",
                         mean="5c.mean",
                         sd="5d.sd",
                         cov="5e.cov",
                         max="5f.max")),
    by=c("key","summ")
  ) %>%
  arrange(key,summ) %>%
  gather(days_from_admit,summ_val,-key,-summ) %>% 
  spread(summ,summ_val)


vital_smoke_summ<-vital %>%
  dplyr::select(PATID,ENCOUNTERID, key, value) %>%
  filter(key %in% c("SMOKING","TOBACCO","TOBACCO_TYPE")) %>%
  unique %>%
  group_by(PATID,ENCOUNTERID, key) %>%
  dplyr::mutate(value=paste(value[order(value)],collapse = ",")) %>% 
  ungroup %>% unique %>%
  spread(key,value) %>%
  right_join(Table1 %>% dplyr::select(PATID,ENCOUNTERID),
             by=c("PATID","ENCOUNTERID")) %>%
  replace_na(list(SMOKING="NI",
                  TOBACCO="NI",
                  TOBACCO_TYPE="NI")) %>%
  gather(key,value,-PATID,-ENCOUNTERID) %>%
  mutate(key2=key) %>%
  unite("key_cat",c("key2","value")) %>%
  group_by(key,key_cat) %>%
  dplyr::summarize(pat_cnt=length(unique(PATID)),
                   enc_cnt=length(unique(ENCOUNTERID)),
                   enc_prop=length(unique(ENCOUNTERID))/enc_tot) %>%
  arrange(desc(pat_cnt)) %>%
  ungroup %>%
  #HIPPA, low counts masking
  mutate(pat_cnt=ifelse(as.numeric(pat_cnt)<11 & as.numeric(pat_cnt)>0,"<11",as.character(pat_cnt)),
         enc_cnt=ifelse(as.numeric(enc_cnt)<11 & as.numeric(enc_cnt)>0,"<11",as.character(enc_cnt)),
         enc_prop=ifelse(as.numeric(enc_cnt)<11 & as.numeric(enc_cnt)>0,"<11",paste0(round(enc_prop,2)*100,"%"))) %>%
  gather(summ,summ_val,-key_cat,-key) %>%
  mutate(summ=recode(summ,
                     pat_cnt="1.patients#",
                     enc_cnt="2.encounters#",
                     enc_prop="3.encounters%")) %>%
  spread(summ,summ_val)


#save
saveRDS(vital,file=paste0("./data/",toupper(sql$tbl_out),".rda"))
# saveRDS(vital_summ,file="./data/vital_summ.rda")
# saveRDS(vital_smoke_summ,file="./data/vital_smoke_summ.rda")

#clean up
rm(vital,vital1); gc()
```


```{r vital_out1, echo=F, warning=F, message=F}
kable(vital_summ %>% dplyr::select(-key),
      caption="Table 2a - Vital (HT,WT,BMI,SBP,DBP) Summaries") %>%
  kable_styling("striped", full_width = F) %>%
  group_rows("BMI", 1,10) %>%
  group_rows("BP_DIASTOLIC", 11, 20) %>%
  group_rows("BP_SYSTOLIC", 21, 30) %>%
  group_rows("HT", 31, 40) %>%  
  group_rows("WT", 41, 50)  
```

Table 2a identifies extreme values of vitals for height, weight, BMI, and blood pressure, which may suggest systemic errors such as typos, and conversion mistakes. Here is the list of how *high* and *low* vitals are defined (adopted from CDM EDC report):

  |vital type  |upper bound for low values|lower bound for high values|     
  |:----------:|:------------------------:|:-------------------------:|   
  |HT(inches)  |0                         |95                         |   
  |WT(lbs)     |0                         |350                        |   
  |BMI         |0                         |50                         |    
  |BP_DIASTOLIC|40                        |120                        |     
  |BP_SYSTOLIC |40                        |210                        |     


```{r vital_out2, echo=F, warning=F, message=F}
vital_smoke_summ %<>% filter(`2.encounters#`!="<11") %>%
  arrange(key,desc(`2.encounters#`))

row_grp_pos<-vital_smoke_summ %>% 
  mutate(rn=1:n()) %>%
  group_by(key) %>%
  dplyr::summarize(begin=rn[1],
                   end=rn[n()]) %>%
  ungroup

kable(vital_smoke_summ %>% dplyr::select(-key),
      caption="Table 2b - Vital (SMOKING, TABACCO) Summaries") %>%
  kable_styling("striped", full_width = F) %>%
  group_rows(row_grp_pos$key[1], row_grp_pos$begin[1],row_grp_pos$end[1]) %>%
  group_rows(row_grp_pos$key[2], row_grp_pos$begin[2],row_grp_pos$end[2]) %>%
  group_rows(row_grp_pos$key[3], row_grp_pos$begin[3],row_grp_pos$end[3])
```


Table 2b identifies unreliable reporting of smoking status. A significant mismatch between smoking and tabacco summaries needs some further investigation.


***


##### Labs
```{r lab, include=F}
## labs
conn<-connect_to_db(params$DBMS_type,config_file)

sql<-parse_sql(paste0("./inst/",params$DBMS_type,"/collect_lab.sql"),
               cdm_db_link=config_file$cdm_db_link,
               cdm_db_name=config_file$cdm_db_name,
               cdm_db_schema=config_file$cdm_db_schema)

lab<-execute_single_sql(conn,
                        statement=sql$statement,
                        write=(sql$action=="write")) %>%
  dplyr::rename(key=LAB_LOINC,value=RESULT_NUM,unit=RESULT_UNIT,
                dsa=DAYS_SINCE_ADMIT,timestamp=SPECIMEN_DATE_TIME) %>%
  dplyr::select(PATID,ENCOUNTERID,key,value,unit,dsa,timestamp) %>%
  filter(!is.na(key) & !is.na(value)) %>%
  unique %>%
  mutate(dsa_grp=case_when(dsa < 0 ~ "0>",
                           dsa >=0 & dsa < 1 ~ "1",
                           dsa >=1 & dsa < 2 ~ "2",
                           dsa >=2 & dsa < 3 ~ "3",
                           dsa >=3 & dsa < 4 ~ "4",
                           dsa >=4 & dsa < 5 ~ "5",
                           dsa >=5 & dsa < 6 ~ "6",
                           dsa >=6 & dsa < 7 ~ "7",
                           dsa >=7 ~ "7<"))  

#collect summaries
lab_summ<-lab %>% 
  group_by(key) %>%
  dplyr::summarize(record_cnt=n(),
                   enc_cnt=length(unique(ENCOUNTERID)),
                   min=min(value,na.rm=T),
                   mean=round(mean(value,na.rm=T),2),
                   sd=round(sd(value,na.rm=T),3),
                   median=round(median(value,na.rm=T)),
                   max=max(value,na.rm=T)) %>%
  ungroup %>%
  mutate(cov=round(sd/mean,3)) %>%
  mutate(freq_rk=rank(-enc_cnt,ties.method="first")) %>%
  #HIPPA, low counts masking
  mutate(enc_cnt=ifelse(as.numeric(enc_cnt)<11 & as.numeric(enc_cnt)>0,"<11",as.character(enc_cnt)),
         record_cnt=ifelse(as.numeric(record_cnt)<11 & as.numeric(record_cnt)>0,"<11",as.character(record_cnt))) %>%
  gather(summ,overall,-key,-freq_rk) %>%
  left_join(
    lab %>%
      group_by(key,dsa_grp) %>%
      dplyr::summarize(record_cnt=n(),
                       enc_cnt=length(unique(ENCOUNTERID)),
                       min=min(value,na.rm=T),
                       mean=round(mean(value,na.rm=T),2),
                       sd=round(sd(value,na.rm=T),3),
                       median=round(median(value,na.rm=T)),
                       max=max(value,na.rm=T)) %>%
      ungroup %>%
      mutate(cov=round(sd/mean,3)) %>%
      #HIPPA, low counts masking
      mutate(enc_cnt=ifelse(as.numeric(enc_cnt)<11 & as.numeric(enc_cnt)>0,"<11",as.character(enc_cnt)),
             record_cnt=ifelse(as.numeric(record_cnt)<11 & as.numeric(record_cnt)>0,"<11",as.character(record_cnt)),
             sd=ifelse(is.nan(sd),0,sd)) %>%
      gather(summ,summ_val,-key,-dsa_grp) %>%
      spread(dsa_grp,summ_val),
    by=c("key","summ")
  ) %>%
  arrange(freq_rk,summ) %>%
  #additional 
  mutate(at_admission=ifelse(is.na(`1`),0,1),
         within_3d=ifelse(is.na(coalesce(`1`,`2`,`3`)),0,1))

#save
saveRDS(lab,file=paste0("./data/",toupper(sql$tbl_out),".rda"))
# saveRDS(lab_summ,file="./data/lab_summ.rda")

#clean up
rm(lab); gc()
```


```{r lab_out, echo=F, warning=F, message=F}
lab_temp<-lab_summ %>%
  filter(summ %in% c("enc_cnt","record_cnt")) %>%
  dplyr::select(key,summ,overall) %>% unique %>%
  filter(overall!="<11") %>%
  mutate(overall=as.numeric(overall)) %>%
  spread(summ,overall,fill=0) %>%
  filter(enc_cnt > 0) %>%
  mutate(record_intensity=round(record_cnt/enc_cnt,2)) %>%
  mutate(label=ifelse(dense_rank(-enc_cnt)<=10 | dense_rank(-record_intensity)<=10,key,""))

ggplot(lab_temp,aes(x=record_intensity,y=enc_cnt,label=label))+
  geom_point()+ geom_text_repel(segment.alpha=0.5,segment.color="grey")+
  scale_y_continuous(sec.axis = sec_axis(trans= ~./enc_tot,
                                         name = 'Percentage'))+
  labs(x="Average Records per Encounter",
       y="Encounter Counts",
       title="Figure 1 - Data Density vs. Records Intensity")

#get their searchable link
lab_report<-lab_temp %>%
  dplyr::filter(key != "NI") %>%
  arrange(desc(enc_cnt)) %>% 
  dplyr::select(key) %>%
  unique %>% slice(1:5) %>%
  bind_rows(
    lab_temp %>% 
      dplyr::filter(key != "NI") %>%
      arrange(desc(record_intensity)) %>% 
      dplyr::select(key) %>%
      unique %>% slice(1:2)
  ) %>%
  mutate(link=lapply(key,get_loinc_ref))

```

A total of `r length(unique(lab_summ$key))` LOINC identifiable labs are eligible (NI may present), among which `r nrow(lab_summ %>% filter(at_admission==1) %>% dplyr::select(key) %>% unique)` are collected at the day of admission, `r nrow(lab_summ %>% filter(within_3d==1) %>% dplyr::select(key) %>% unique)` within 3 days. Figure 1 shows the data density and intensity of labs concepts, which can help identify common labs (e.g. the common labs for this study cohort are [`r lab_report$key[1]`], [`r lab_report$key[2]`],[`r lab_report$key[3]`],[`r lab_report$key[4]`],[`r lab_report$key[5]`],...), and labs with very high recording intensity (e.g. [`r lab_report$key[6]`], [`r lab_report$key[7]`]).

[`r lab_report$key[1]`]: `r lab_report$link[1]`
[`r lab_report$key[2]`]: `r lab_report$link[2]` 
[`r lab_report$key[3]`]: `r lab_report$link[3]`
[`r lab_report$key[4]`]: `r lab_report$link[4]`
[`r lab_report$key[5]`]: `r lab_report$link[5]`
[`r lab_report$key[6]`]: `r lab_report$link[6]`
[`r lab_report$key[7]`]: `r lab_report$link[7]`


***


##### Diagnosis
```{r dx, include=F}
load("./data/ccs_icd_cw.Rdata")
conn<-connect_to_db(params$DBMS_type,config_file)

## historical diagnosis
sql<-parse_sql(paste0("./inst/",params$DBMS_type,"/collect_dx.sql"),
               cdm_db_link=config_file$cdm_db_link,
               cdm_db_name=config_file$cdm_db_name,
               cdm_db_schema=config_file$cdm_db_schema)

dx<-execute_single_sql(conn,
                       statement=sql$statement,
                       write=(sql$action=="write")) %>%
  #attach CCS diagnosis grouping
  dplyr::mutate(DX_ICD=paste0("ICD",DX_TYPE,":",DX)) %>%
  left_join(ccs_icd %>% select(-ccs_name),by=c("DX_ICD"="icd_w_type")) %>%
  unique %>% filter(!is.na(ccs_code)) %>%
  dplyr::rename(key=ccs_code, dsa=DAYS_SINCE_ADMIT) %>%
  dplyr::select(PATID,ENCOUNTERID,key,dsa) %>%
  unique

#collect summaries
dx_summ<-dx %>%
  group_by(key) %>%
  dplyr::summarize(record_cnt=n(),
                   pat_cnt=length(unique(PATID)),
                   enc_cnt=length(unique(ENCOUNTERID)),
                   min_history=min(dsa,na.rm=T),
                   mean_history=round(mean(dsa,na.rm=T)),
                   sd_history=round(sd(dsa,na.rm=T)),
                   median_history=round(median(dsa,na.rm=T)),
                   max_history=max(dsa,na.rm=T)) %>%
  ungroup %>%
  #HIPPA, low counts masking
  mutate(pat_cnt=ifelse(as.numeric(pat_cnt)<11,"<11",pat_cnt),
         enc_cnt=ifelse(as.numeric(enc_cnt)<11,"<11",enc_cnt),
         record_cnt=ifelse(as.numeric(record_cnt)<11,"<11",record_cnt)) %>%
  arrange(key)

#save
saveRDS(dx,file=paste0("./data/",toupper(sql$tbl_out),".rda"))
# saveRDS(dx_summ,file="./data/dx_summ.rda")

#clean up
rm(dx); gc()
```



```{r dx_out, echo=F, warning=F, message=F}
load("./data/ccs_ref.Rdata")

#historical diagnosis
dx_temp<-dx_summ %>% 
  dplyr::select(key,enc_cnt,mean_history) %>%
  filter(enc_cnt!="<11") %>%
  mutate(enc_cnt=as.numeric(enc_cnt),
         mean_history=as.numeric(mean_history)) %>%
  filter(enc_cnt > 0) %>%
  mutate(label=ifelse(dense_rank(-enc_cnt)<=10,key,""))

ggplot(dx_temp,aes(x=mean_history,y=enc_cnt,label=label))+
  geom_point()+geom_text_repel()+
  scale_y_continuous(sec.axis = sec_axis(trans= ~./enc_tot,
                                         name = 'Percentage'))+
  labs(x="Mean History of Diagnoses (Days)",
       y="Encounter Counts",
       title="Figure 2 - Data Density vs. Recency (CCS)")

dx_report<-dx_temp %>% 
  arrange(desc(enc_cnt)) %>%
  slice(1:6) %>%
  dplyr::select(key) %>%
  left_join(ccs_ref %>% filter(type=="dx"),
            by=c("key"="ccs_code")) %>%
  dplyr::select(key,ccs_name)
```


A Total of `r length(unique(dx_summ$key))` distinct CCS-grouped diagnoses has been assigned to patients before the encounter of interest. Figure 2 gives an overview of average history of patients' diagnosis prior to tne encounter of interest as well as the highly frequent historical diagnoses(e.g. `r dx_report$key[1]`(`r dx_report$ccs_name[1]`), `r dx_report$key[2]`(`r dx_report$ccs_name[2]`), `r dx_report$key[3]`(`r dx_report$ccs_name[3]`), `r dx_report$key[4]`(`r dx_report$ccs_name[4]`), `r dx_report$key[5]`(`r dx_report$ccs_name[5]`), `r dx_report$key[6]`(`r dx_report$ccs_name[6]`)).


***


##### Procedure
```{r px, include=F}
## procedure
conn<-connect_to_db(params$DBMS_type,config_file)

sql<-parse_sql(paste0("./inst/",params$DBMS_type,"/collect_px.sql"),
               cdm_db_link=config_file$cdm_db_link,
               cdm_db_name=config_file$cdm_db_name,
               cdm_db_schema=config_file$cdm_db_schema)

px<-execute_single_sql(conn,
                       statement=sql$statement,
                       write=(sql$action=="write")) %>%
  dplyr::mutate(PX=paste0(PX_TYPE,":",PX)) %>%
  dplyr::rename(key=PX,dsa=DAYS_SINCE_ADMIT) %>%
  dplyr::select(PATID,ENCOUNTERID,key,dsa) %>%
  unique %>%
  mutate(dsa_grp=case_when(dsa < 0 ~ "0>",
                           dsa >=0 & dsa < 1 ~ "1",
                           dsa >=1 & dsa < 2 ~ "2",
                           dsa >=2 & dsa < 3 ~ "3",
                           dsa >=3 & dsa < 4 ~ "4",
                           dsa >=4 & dsa < 5 ~ "5",
                           dsa >=5 & dsa < 6 ~ "6",
                           dsa >=6 & dsa < 7 ~ "7",
                           dsa >=7 ~ "7<"))  

px_summ<-px %>%
  group_by(key,dsa_grp) %>%
  dplyr::summarize(record_cnt=n(),
                   pat_cnt=length(unique(PATID)),
                   enc_cnt=length(unique(ENCOUNTERID))) %>%
  ungroup %>%
  #HIPPA, low counts masking
  mutate(pat_cnt=ifelse(as.numeric(pat_cnt)<11,"<11",pat_cnt),
         enc_cnt=ifelse(as.numeric(enc_cnt)<11,"<11",enc_cnt),
         record_cnt=ifelse(as.numeric(record_cnt)<11,"<11",record_cnt)) %>%
  arrange(key,dsa_grp)

#save
saveRDS(px,file=paste0("./data/",toupper(sql$tbl_out),".rda"))
# saveRDS(px_summ,file="./data/px_summ.rda")

#clean up
rm(px); gc()
```


```{r px_out, echo=F, warning=F, message=F}
px_temp<-px_summ %>% 
  dplyr::select(key,dsa_grp,enc_cnt) %>% 
  filter(enc_cnt!="<11") %>%
  mutate(enc_cnt=as.numeric(enc_cnt)) %>%
  filter(enc_cnt > 0) %>%
  arrange(desc(enc_cnt)) %>%
  mutate(label=ifelse(dense_rank(-enc_cnt)<=10,key,""))

ggplot(px_temp,aes(x=dsa_grp,y=enc_cnt,label=label))+
  geom_point()+geom_text_repel()+
  scale_y_continuous(sec.axis = sec_axis(trans= ~./enc_tot,
                                         name = 'Percentage'))+
  labs(x="Days since Admission",
       y="Encounter Counts",
       title="Figure 3 - Procedure Density over Time")

px_report<-px_temp %>%
  arrange(desc(enc_cnt)) %>%
  dplyr::select(key) %>%
  unique %>% slice(1:5) %>%
  # mutate(link=lapply(key,google_code)) %>%
  mutate(link="https://coder.aapc.com/cpt-codes/")
  
```


A Total of `r length(unique(px_summ$key))` distinct total procedures codes have been assigned to patients before or at the encounter of interest. Figure3 gives an overview of procedures that patients had recieved before or at the encounters of interest. It can help identify the common procedures or typical occuring times of precedures (e.g. [`r px_report$key[1]`], [`r px_report$key[2]`],[`r px_report$key[3]`],[`r px_report$key[4]`],[`r px_report$key[5]`],...). Note that Figure2 and Figure3 may display similar distributions as a result of corrlations between diagnoses and procedures.

[`r px_report$key[1]`]: `r px_report$link[1]`
[`r px_report$key[2]`]: `r px_report$link[2]` 
[`r px_report$key[3]`]: `r px_report$link[3]`
[`r px_report$key[4]`]: `r px_report$link[4]`
[`r px_report$key[5]`]: `r px_report$link[5]`

*note: links for cpt codes may lead to invalid page* 

***


##### Medications
```{r med, include=F}
## medication
conn<-connect_to_db(params$DBMS_type,config_file)

sql<-parse_sql(paste0("./inst/",params$DBMS_type,"/collect_med",
                      ifelse(params$incl_NDC,"_ndc",""),".sql"),
               cdm_db_link=config_file$cdm_db_link,
               cdm_db_name=config_file$cdm_db_name,
               cdm_db_schema=config_file$cdm_db_schema)

med<-execute_single_sql(conn,
                        statement=sql$statement,
                        write=(sql$action=="write")) %>%
  dplyr::mutate(RX_EXPOS=round(pmin(pmax(as.numeric(difftime(RX_END_DATE,RX_START_DATE,units="days")),1),
                                    pmax(RX_DAYS_SUPPLY,1),na.rm=T))) %>%
  replace_na(list(RX_QUANTITY_DAILY=1)) %>%
  dplyr::rename(sdsa=DAYS_SINCE_ADMIT) %>%
  mutate(RX_ID=ifelse(params$incl_NDC,paste(RXNORM_CUI,RAW_RX_NDC,collapse="_"),RXNORM_CUI)) %>%
  dplyr::select(PATID,ENCOUNTERID,RX_ID,RX_BASIS,RX_EXPOS,RX_QUANTITY_DAILY,sdsa) %>%
  unite("key",c("RX_ID","RX_BASIS"),sep=":")

#save intermediate table for debugging
saveRDS(med,file=paste0("./data/",toupper(sql$tbl_out),".rda"))

#converted to daily exposure
batch<-20
expos_quant<-c(1,unique(quantile(med[med$RX_EXPOS>1,]$RX_EXPOS,probs=0:batch/batch)))
med2<-med %>% filter(RX_EXPOS<=1) %>% 
  dplyr::mutate(dsa=as.character(sdsa),edsa=sdsa,value=RX_QUANTITY_DAILY) %>%
  dplyr::select(PATID,ENCOUNTERID,key,value,sdsa,edsa,dsa)

for(i in seq_len(length(expos_quant)-1)){
  med_sub<-med %>% filter(RX_EXPOS > expos_quant[i] & RX_EXPOS <= expos_quant[i+1])
  med_expand<-med_sub[rep(row.names(med_sub),(med_sub$RX_EXPOS+1)),] %>%
    group_by(PATID,ENCOUNTERID,key,RX_QUANTITY_DAILY,sdsa) %>%
    dplyr::mutate(expos_daily=1:n()-1) %>% 
    dplyr::summarize(edsa=max(sdsa+expos_daily),
                     dsa=paste0(sdsa+expos_daily,collapse=",")) %>%
    ungroup %>% dplyr::rename(value=RX_QUANTITY_DAILY) %>%
    dplyr::select(PATID,ENCOUNTERID,key,value,sdsa,edsa,dsa)
  med2 %<>% bind_rows(med_expand)
  gc()
}

#merge overlapped precribing intervals
med2 %<>% 
  group_by(PATID,ENCOUNTERID,key,sdsa,edsa,dsa) %>%
  dplyr::summarize(value=value[1]) %>%
  ungroup

med<-med2 %>%
  group_by(PATID,ENCOUNTERID,key,sdsa) %>%
  dplyr::summarize(RX_EXPOS=pmax(1,sum(value,na.rm=T))) %>%
  ungroup


#collect summaries
if(params$incl_NDC){
    med_summ<-med %>%
    mutate(dsa_grp=case_when(sdsa < 0 ~ "0>",
                             sdsa >=0 & sdsa < 7 ~ "1w",
                             sdsa >=7 ~ "7<")) 
}else{
  med_summ<-med %>%
    mutate(dsa_grp=case_when(sdsa < 0 ~ "0>",
                             sdsa >=0 & sdsa < 1 ~ "1",
                             sdsa >=1 & sdsa < 2 ~ "2",
                             sdsa >=2 & sdsa < 3 ~ "3",
                             sdsa >=3 & sdsa < 4 ~ "4",
                             sdsa >=4 & sdsa < 5 ~ "5",
                             sdsa >=5 & sdsa < 6 ~ "6",
                             sdsa >=6 & sdsa < 7 ~ "7",
                             sdsa >=7 ~ "7<")) 
}

med_summ %<>%
  group_by(key,dsa_grp) %>%
  dplyr::summarize(record_cnt=n(),
                   enc_cnt=length(unique(ENCOUNTERID)),
                   min_expos=min(RX_EXPOS,na.rm=T),
                   mean_expos=round(mean(RX_EXPOS,na.rm=T)),
                   sd_expos=round(sd(RX_EXPOS,na.rm=T)),
                   median_expos=round(median(RX_EXPOS,na.rm=T)),
                   max_expos=max(RX_EXPOS,na.rm=T)) %>%
  ungroup %>%
  #HIPPA, low counts masking
  mutate(enc_cnt=ifelse(as.numeric(enc_cnt)<11,"<11",as.character(enc_cnt)),
         record_cnt=ifelse(as.numeric(record_cnt)<11,"<11",as.character(record_cnt)),
         sd_expos=ifelse(is.na(sd_expos),0,sd_expos)) %>%
  dplyr::mutate(cov_expos=round(sd_expos/mean_expos,1)) %>%
  gather(summ,summ_val,-key,-dsa_grp) %>%
  spread(dsa_grp,summ_val) %>%
  arrange(key,summ) 

med_density<-length(unique(ENCOUNTERID))

#save
saveRDS(med2,file=paste0("./data/",toupper(sql$tbl_out),".rda"))
# saveRDS(med_summ,file="./data/med_summ.rda")

#clean up
rm(med,med2); gc()
```


```{r med_out, echo=F, warning=F, message=F, fig.height = 8, fig.width = 12}
med_temp<-med_summ %>% 
  filter(summ %in% c("enc_cnt","median_expos")) %>% 
  gather(dsa_grp,summ_val,-summ,-key) %>%
  filter(!is.na(summ_val) & (summ_val!="<11")) %>%
  mutate(summ_val=as.numeric(summ_val)) %>%
  spread(summ,summ_val) %>%
  filter(!is.na(median_expos) & enc_cnt>=enc_tot*0.001) %>%
  arrange(median_expos) %>%
  group_by(dsa_grp) %>%
  dplyr::mutate(label=ifelse(dense_rank(-enc_cnt)<=3,key,"")) %>%
  ungroup %>%
  dplyr::mutate(label=ifelse(label!="",label,
                             ifelse(dense_rank(-median_expos)<=2,key,"")))


if(nrow(med_temp)>0){
  ggplot(med_temp,aes(x=dsa_grp,y=enc_cnt,color=median_expos,label=label)) +
    geom_point() + geom_text_repel()+
    scale_y_continuous(sec.axis = sec_axis(trans= ~./enc_tot,
                                           name = 'Percentage'))+
    scale_color_gradient2(low = "green",mid="blue",high ="red",
                          midpoint = 2)+
    labs(x="Start Date",y="Encounter Counts",color="Median Exposure (days)",
         title="Figure 4 - Medication Exposure Summaries")
  
  med_report<-med_temp %>%
    mutate(key=gsub(".*_","",gsub("\\:.*","",key))) %>%
    arrange(desc(enc_cnt)) %>%
    dplyr::select(key) %>% 
    unique %>% slice(1:3) %>%
    mutate(rx_name=lapply(key,function(x) ifelse(params$incl_NDC,
                                                 get_ndc_nm(x),
                                                 get_rxcui_nm(x))))
  
  freq_med<-c()
  for(k in 1:nrow(med_report)){
    freq_med<-c(freq_med,paste0(med_report$key[k],"(",med_report$rx_name[k],")")) 
  }
  
  med_report<-med_temp %>%
    mutate(key=gsub(".*_","",gsub("\\:.*","",key))) %>%
    arrange(desc(median_expos)) %>%
    dplyr::select(key) %>%
    unique %>% slice(1:3) %>%
    mutate(rx_name=lapply(key,function(x) ifelse(params$incl_NDC,
                                                 get_ndc_nm(x),
                                                 get_rxcui_nm(x))))
  
  intens_med<-c()
  for(k in 1:nrow(med_report)){
    intens_med<-c(intens_med,paste0(med_report$key[k],"(",med_report$rx_name[k],")")) 
  }
  
  description<-paste0("Figure4 demonstrates average exposures for drug starting at X days since admission. 
                      It helps identify typical medciations dispensed (:01) or administered(:02) during the course of stay. 
                      (e.g. the typical medications identified are",paste(freq_med,collapse=","),
                      "; while drugs such as ",paste(intens_med,collapse=","), 
                      "are used with a relative longer exposure than the others).")
}else{
  description<-"Medication exposure are too low as no medication identifier has a coverage of more than 0.1% of the study population."
}

```

A Total of `r length(unique(med_summ$key))` distinct `r ifelse(params$incl_NDC,"NDC","RXNORM")` medication concepts are discovered for the cohort and the overall medication (any) exposure for this cohort is `r paste0(round(med_density/enc_tot,2)*100,"%")`. `r description`


```{r final, include=F}
final_out<-list(Table1=consort_tbl,
                Table2=tbl1_summ,
                Table3=demo_nice_tbl,
                Table4=vital_summ,
                Table5=vital_smoke_summ,
                Table6=lab_summ,
                Table7=dx_summ,
                Table8=px_summ,
                Table9=med_summ)
write.xlsx(final_out,file="./output/AKI_CDM_EXT_VALID_p1_QA_TBL.xlsx")


# ggplot sometimes create some unwanted empty .pdf file and want to clean it up
if(file.exists("./Rplots.pdf")){
  file.remove("./Rplots.pdf")
}

rm(list=ls())
gc()
```

***


