---
title: "Building and Validating Predictive Models for Acute Kidney Injury (AKI) using PCORnet CDM (Part II.1)"
author: "xing song"
date: "Feburary 09, 2019"
output: html_document
---
### Stage 2: Predictive Models Validation (Benchmark)   

In this experiment, we will build the benchmark predictive model by quasi-replicating the model by [*Koyner et al*] for AKI risk prediction on the adult inpatients at each GPC site using PCORnet common data model. This learning frame work by Koyner et al is, by far, the best reported general inpatients AKI prediction model, which is based on a tuned gradient boosting machine in a discrete time survival analysis framework with pre-selected EHR data elements. We aim to extend this benchmark prediction model by including novel elements from PCORnet CDM, while effectively automate the manual efforts of feature pre-selection and cleaning. The primary outcomes of this study were the development of AKI stage ≥ 1, AKI stage ≥ 2, and AKI stage 3 in the next 24 and 48 hours.

[*Koyner et al*] https://www.ncbi.nlm.nih.gov/pubmed/29596073 

***

```{r setup, include=F}
#source utility functions
source("./R/util.R")
source("./R/var_etl_surv.R")

require_libraries(c("tidyr",
                    "dplyr",
                    "magrittr",
                    "stringr",
                    "broom",
                    "Matrix",
                    "xgboost",
                    "ROCR",
                    "PRROC"))

```


#### Preparation    

By running both parts of "render_report.R", the raw data tables should have already been collected and saved in the local `./data` folder (Note: these data tables are not visiable in the github ./data folder), that are
* `Table1.rda`: AKI patieht cohort table;        
* `AKI_DEMO.rda`: CDM demographic table cut for AKI cohort;   
* `AKI_VITAL.rda`: CDM vital table cut for AKI cohort;    
* `AKI_LAB.rda`: CDM lab table cut for AKI cohort;
* `AKI_DX.rda`: CDM diagnosis table cut for AKI cohort;   
* `AKI_PX.rda`: CDM procedure table cut for AKI cohort;   
* `AKI_MED.rda`: CDM prescribing medication table cut for AKI cohort;   

***

#### Objective 2.1: Data Cleaning and Representation

In this section, the raw data tables will be cleaned and transformed to a discrete-survival-like representation, which will be used in the final modeling stage. To reduce the burden on memory requirments, the ETL (extract, transform, load) process will be performed in chunks with respect to **distinct prediction task, encounter years and variable types**. Meanwhile, indices for random paritioning will be assigned to each encounter. The ETL progress will be reported as follows:


```{r preprocess, include=F}
# experimental design parameters
#-----prediction point
pred_in_d_opt<-c(1,2)
#-----prediction tasks
pred_task_lst<-c("stg1up","stg2up","stg3")

############################## collect and format variables on daily basis ######################
tbl1<-readRDS("./data//Table1.rda") %>%
  dplyr::mutate(yr=as.numeric(format(strptime(ADMIT_DATE, "%Y-%m-%d %H:%M:%S"),"%Y")))

for(pred_in_d in pred_in_d_opt){
  #--determine update time window
  tw<-as.double(seq(0,7))
  if(pred_in_d>1){
  tw<-tw[-seq_len(pred_in_d-1)]
  } 
  
  #--by chunks: encounter year
enc_yr<-tbl1 %>%
  dplyr::select(yr) %>%
  unique %>% arrange(yr) %>%
  filter(yr>2009) %>%
  unlist

#--by variable type
var_type<-c("demo","vital","lab","dx","px","med")

#--save results as array
for(pred_task in pred_task_lst){
  start_tsk<-Sys.time()
  cat("Start variable collection for task",pred_task,".\n")
  #---------------------------------------------------------------------------------------------
  
  var_by_yr<-list()
  var_bm<-list()
  rsample_idx<-c()
  for(i in seq_along(enc_yr)){
    start_i<-Sys.time()
    cat("...start variable collection for year",enc_yr[i],".\n")
    
    #--collect end_points
    dat_i<-tbl1 %>% filter(yr==enc_yr[i]) %>%
      dplyr::select(ENCOUNTERID,yr,
                    NONAKI_SINCE_ADMIT,
                    AKI1_SINCE_ADMIT,
                    AKI2_SINCE_ADMIT,
                    AKI3_SINCE_ADMIT) %>%
      gather(y,dsa_y,-ENCOUNTERID,-yr) %>%
      filter(!is.na(dsa_y)) %>%
      dplyr::mutate(y=recode(y,
                             "NONAKI_SINCE_ADMIT"=0,
                             "AKI1_SINCE_ADMIT"=1,
                             "AKI2_SINCE_ADMIT"=2,
                             "AKI3_SINCE_ADMIT"=3)) %>%
      dplyr::mutate(y=as.numeric(y))
    
    if(pred_task=="stg1up"){
      dat_i %<>%
        dplyr::mutate(y=as.numeric(y>0)) %>%
        group_by(ENCOUNTERID) %>% top_n(n=1L,wt=dsa_y) %>% ungroup
    }else if(pred_task=="stg2up"){
      dat_i %<>%
        # filter(y!=1) %>% # remove stage 1
        dplyr::mutate(y=as.numeric(y>1)) %>%
        group_by(ENCOUNTERID) %>% top_n(n=1L,wt=dsa_y) %>% ungroup
    }else if(pred_task=="stg3"){
      dat_i %<>%
        # filter(!(y %in% c(1,2))) %>% # remove stage 1,2
        dplyr::mutate(y=as.numeric(y>2)) %>%
        group_by(ENCOUNTERID) %>% top_n(n=1L,wt=dsa_y) %>% ungroup
    }else{
      stop("prediction task is not valid!")
    }
    
    #--random sampling
    rsample_idx %<>%
      bind_rows(dat_i %>% 
                  dplyr::select(ENCOUNTERID,yr) %>%
                  unique %>%
                  dplyr::mutate(cv10_idx=sample(1:10,n(),replace=T)))
    
    #--ETL variables
    X_surv<-c()
    y_surv<-c()
    var_etl_bm<-c()
    for(v in seq_along(var_type)){
      start_v<-Sys.time()
      
      #extract
      var_v<-readRDS(paste0("./data_local/data_raw/AKI_",toupper(var_type[v]),".rda")) %>%
        semi_join(dat_i,by="ENCOUNTERID")
      
      if(var_type[v] != "demo"){
        if(var_type[v] == "med"){
          var_v %<>% 
            transform(value=strsplit(value,","),
                      dsa=strsplit(dsa,",")) %>%
            unnest(value,dsa) %>%
            dplyr::mutate(value=as.numeric(value),
                          dsa=as.numeric(dsa))
        }
        var_v %<>% filter(dsa <= pred_end)
      }
      
      #transform
      var_v<-format_data(dat=var_v,
                         type=var_type[v],
                         pred_end=pred_end)
      
      Xy_surv<-get_dsurv_temporal(dat=var_v,
                                  censor=dat_i,
                                  tw=tw,
                                  pred_in_d=pred_in_d)
      
      #load
      X_surv %<>% bind_rows(Xy_surv$X_surv) %>% unique
      y_surv %<>% bind_rows(Xy_surv$y_surv) %>% unique
      
      lapse_v<-Sys.time()-start_v
      var_etl_bm<-c(var_etl_bm,paste0(lapse_v,units(lapse_v)))
      cat("\n......finished ETL",var_type[v],"for year",enc_yr[i],"in",lapse_v,units(lapse_v),".\n")
    }
    var_by_yr[[i]]<-list(X_surv=X_surv,
                         y_surv=y_surv)
    
    lapse_i<-Sys.time()-start_i
    var_etl_bm<-c(var_etl_bm,paste0(lapse_i,units(lapse_i)))
    cat("\n...finished variabl collection for year",enc_yr[i],"in",lapse_i,units(lapse_i),".\n")
    
    var_bm[[i]]<-data.frame(bm_nm=c(var_type,"overall"),
                            bm_time=var_etl_bm,
                            stringsAsFactors = F)
  }
  #--save preprocessed data
  saveRDS(rsample_idx,file=paste0("./data/",pred_in_d,"d_rsample_idx_",pred_task,".rda"))
  saveRDS(var_by_yr,file=paste0("./data/",pred_in_d,"d_var_by_yr_",pred_task,".rda"))
  saveRDS(var_bm,file=paste0("./data/",pred_in_d,"d_var_bm",pred_task,".rda"))
  
  #---------------------------------------------------------------------------------------------
  lapse_tsk<-Sys.time()-start_tsk
  cat("\nFinish variable ETL for task:",pred_task,"in",pred_in_d,"days",",in",lapse_tsk,units(lapse_tsk),".\n")
  }
}
```

The final outputs from this code chunk should be found in the `./data` folder as the following intermediate data tables for different prediction tasks:
* For AKI stage ≥ 1 in 24 hours: `1d_rsample_idx_stg1up.rda`, `1d_var_by_yr_stg1up.rda`, `1d_var_bm_stg1up.rda`
* For AKI stage ≥ 2 in 24 hours: `1d_rsample_idx_stg2up.rda`, `1d_var_by_yr_stg2up.rda`, `1d_var_bm_stg2up.rda`
* For AKI stage = 3 in 24 hours: `1d_rsample_idx_stg3.rda`, `1d_var_by_yr_stg3.rda`, `1d_var_bm_stg3.rda`
* For AKI stage ≥ 1 in 48 hours: `2d_rsample_idx_stg1up.rda`, `2d_var_by_yr_stg1up.rda`, `1d_var_bm_stg1up.rda`
* For AKI stage ≥ 2 in 48 hours: `2d_rsample_idx_stg2up.rda`, `2d_var_by_yr_stg2up.rda`, `1d_var_bm_stg2up.rda`
* For AKI stage = 3 in 48 hours: `2d_rsample_idx_stg3.rda`, `2d_var_by_yr_stg3.rda`, `1d_var_bm_stg3.rda`


***

#### Objective 2.2: Benchmark Model Development

We adopted the AKI prediction model by [*Koyner et al*] using 


```{r benchmark,include=F}
#-----feature selection type
fs_type_opt<-c("no_fs","rm_scr_bun")
rm_key<-c('2160-0','38483-4','14682-9','21232-4','35203-9','44784-7','59826-8',
          '16188-5','16189-3','59826-8','35591-7','50380-5','50381-3','35592-5',
          '44784-7','11041-1','51620-3','72271-0','11042-9','51619-5','35203-9','14682-9',
          '12966-8','12965-0','6299-2','59570-2','12964-3','49071-4','72270-2',
          '11065-0','3094-0','35234-4','14937-7',
          '3097-3','44734-2','BUN_SCR')

for(pred_task in pred_task_lst){
  bm<-c()
  bm_nm<-c()
  
  start_tsk<-Sys.time()
  # cat("Start build reference model for task",pred_task,".\n")
  #---------------------------------------------------------------------------------------------
  
  start_tsk_i<-Sys.time()
  #--prepare training and testing set
  yr_rg<-seq(2010,2018)
  X_tr<-c()
  X_ts<-c()
  y_tr<-c()
  y_ts<-c()
  rsample_idx<-readRDS(paste0("./data_local/data_preproc/",pred_in_d,"d_rsample_idx_",pred_task,".rda"))
  for(i in seq_along(yr_rg)){
    var_by_yr<-readRDS(paste0("./data_local/data_preproc/",pred_in_d,"d_var_by_yr_",pred_task,".rda"))[[i]]
    
    X_tr %<>% bind_rows(var_by_yr[["X_surv"]]) %>%
      semi_join(rsample_idx %>% filter(cv10_idx<=6 & yr<2017),
                by="ENCOUNTERID")
    
    y_tr %<>% bind_rows(var_by_yr[["y_surv"]] %>%
                          left_join(rsample_idx %>% filter(cv10_idx<=6 & yr<2017),
                                    by="ENCOUNTERID"))
    
    X_ts %<>% bind_rows(var_by_yr[["X_surv"]]) %>%
      semi_join(rsample_idx %>% filter(cv10_idx>6 | yr>=2017),
                by="ENCOUNTERID")
    
    y_ts %<>% bind_rows(var_by_yr[["y_surv"]] %>%
                          left_join(rsample_idx %>% filter(cv10_idx>6 | yr>=2017),
                                    by="ENCOUNTERID"))
    
    # cat("...finish stack data of encounters from",yr_rg[i],".\n")
  }
  lapse_i<-Sys.time()-start_tsk_i
  bm<-c(bm,paste0(lapse_i,units(lapse_i)))
  bm_nm<-c(bm_nm,"prepare data")
  
  for(fs_type in fs_type_opt){
      start_tsk_i<-Sys.time()
  #--pre-filter
  if(fs_type=="rm_scr_bun"){
    X_tr %<>%
      filter(!(key %in% c(rm_key,paste0(rm_key,"_change"))))
    
    X_ts %<>%
      filter(!(key %in% c(rm_key,paste0(rm_key,"_change"))))
  }
  
  #--transform training matrix
  y_tr %<>%
    filter(!is.na(cv10_idx)) %>%
    arrange(ENCOUNTERID,dsa_y) %>%
    unite("ROW_ID",c("ENCOUNTERID","dsa_y")) %>%
    arrange(ROW_ID) %>%
    unique
  
  X_tr %<>%
    arrange(ENCOUNTERID,dsa_y) %>%
    unite("ROW_ID",c("ENCOUNTERID","dsa_y")) %>%
    semi_join(y_tr,by="ROW_ID") %>%
    long_to_sparse_matrix(df=.,
                          id="ROW_ID",
                          variable="key",
                          val="value")
  
  #--collect variables used in training
  tr_key<-data.frame(key = unique(colnames(X_tr)),
                     stringsAsFactors = F)
  
  #--transform testing matrix
  y_ts %<>%
    filter(!is.na(cv10_idx)) %>%
    arrange(ENCOUNTERID,dsa_y) %>%
    unite("ROW_ID",c("ENCOUNTERID","dsa_y")) %>%
    arrange(ROW_ID) %>%
    unique
  
  X_ts %<>% 
    unite("ROW_ID",c("ENCOUNTERID","dsa_y")) %>%
    semi_join(y_ts,by="ROW_ID") %>%
    semi_join(tr_key,by="key")
  
  x_add<-tr_key %>%
    anti_join(data.frame(key = unique(X_ts$key),
                         stringsAsFactors = F),
              by="key")
  
  #align with training
  if(nrow(x_add)>0){
    X_ts %<>%
      arrange(ROW_ID) %>%
      bind_rows(data.frame(ROW_ID = rep("0_0",nrow(x_add)),
                           dsa = -99,
                           key = x_add$key,
                           value = 0,
                           stringsAsFactors=F))
  }
  X_ts %<>%
    long_to_sparse_matrix(df=.,
                          id="ROW_ID",
                          variable="key",
                          val="value")
  if(nrow(x_add)>0){
    X_ts<-X_ts[-1,]
  }
  
  #check alignment
  all(row.names(X_tr)==y_tr$ROW_ID)
  all(row.names(X_ts)==y_ts$ROW_ID)
  all(colnames(X_tr)==colnames(X_ts))
  
  #--covert to xgb data frame
  dtrain<-xgb.DMatrix(data=X_tr,label=y_tr$y)
  dtest<-xgb.DMatrix(data=X_ts,label=y_ts$y)
  
  lapse_i<-Sys.time()-start_tsk_i
  bm<-c(bm,paste0(lapse_i,units(lapse_i)))
  bm_nm<-c(bm_nm,"transform data")
  
  # cat("...finish formatting training and testing sets.\n")
  
  start_tsk_i<-Sys.time()
  #--get indices for k folds
  y_tr %<>% dplyr::mutate(row_idx = 1:n())
  folds<-list()
  for(fd in seq_len(max(y_tr$cv10_idx))){
    fd_df<-y_tr %>% 
      filter(cv10_idx==fd) %>%
      dplyr::select(row_idx)
    folds[[fd]]<-fd_df$row_idx
  }
  
  #--tune hyperparameter
  #hyper-parameter grid for xgboost
  eval_metric<-"auc"
  objective<-"binary:logistic"
  grid_params<-expand.grid(
    max_depth=c(4,6,10),
    eta=c(0.3,0.1,0.01),
    min_child_weight=1,
    subsample=0.8,
    colsample_bytree=0.8, 
    gamma=1
  )
  
  verb<-TRUE
  bst_grid<-c()
  bst_grid_cv<-c()
  metric_name<-paste0("test_", eval_metric,"_mean")
  metric_sd_name<-paste0("test_", eval_metric,"_std")
  
  for(i in seq_len(dim(grid_params)[1])){
    start_i<-Sys.time()
    param<-as.list(grid_params[i,])
    # param$scale_pos_weight=mean(train$y_train$DKD_IND_additive) #inbalance sampling
    param$scale_pos_weight=1 #balance sampling
    
    bst <- xgb.cv(param,
                  dtrain,
                  objective = objective,
                  metrics = eval_metric,
                  maximize = TRUE,
                  nrounds=2000,
                  # nfold = 5,
                  folds = folds,
                  early_stopping_rounds = 100,
                  print_every_n = 100,
                  prediction = T) #keep cv results
    
    bst_grid<-rbind(bst_grid, cbind(grid_params[i,],
                                    metric=max(bst$evaluation_log[[metric_name]]),
                                    steps=which(bst$evaluation_log[[metric_name]]==max(bst$evaluation_log[[metric_name]]))[1]))
    
    bst_grid_cv<-cbind(bst_grid_cv,bst$pred)
    
    if(verb){
      cat('...finished train case:',paste0(paste0(c(colnames(grid_params),"scale_pos_weight"),"="),param,collapse="; "),
          'in',Sys.time()-start_i,units(Sys.time()-start_i),"\n")
      start_i<-Sys.time()
    }
  }
  hyper_param<-bst_grid[which.max(bst_grid$metric),]
  
  lapse_i<-Sys.time()-start_tsk_i
  bm<-c(bm,paste0(lapse_i,units(lapse_i)))
  bm_nm<-c(bm_nm,"tune model")
  
  # cat("...finish model tunning.\n")

  start_tsk_i<-Sys.time()  
  #--validation
  xgb_tune<-xgb.train(data=dtrain,
                      max_depth=hyper_param$max_depth,
                      maximize = TRUE,
                      eta=hyper_param$eta,
                      nrounds=hyper_param$steps,
                      eval_metric="auc",
                      objective="binary:logistic",
                      print_every_n = 100)
  
  valid<-data.frame(y_ts,
                    pred = predict(xgb_tune,dtest),
                    stringsAsFactors = F)
  
  #--feature importance
  feat_imp<-xgb.importance(colnames(X_tr),model=xgb_tune)
  
  lapse_i<-Sys.time()-start_tsk_i
  bm<-c(bm,paste0(lapse_i,units(lapse_i)))
  bm_nm<-c(bm_nm,"validate model")
  
  # cat("...finish model validating.\n")
  
  #--save model and other results
  saveRDS(xgb_tune,file=paste0("./data_local/model_ref/pred_in_",pred_in_d,"d_model_gbm_",fs_type,"_",pred_task,".rda"))
  saveRDS(bst_grid,file=paste0("./data_local/model_ref/pred_in_",pred_in_d,"d_hyperpar_gbm_",fs_type,"_",pred_task,".rda"))
  saveRDS(valid,file=paste0("./data_local/model_ref/pred_in_",pred_in_d,"d_valid_gbm_",fs_type,"_",pred_task,".rda"))
  saveRDS(feat_imp,file=paste0("./data_local/model_ref/pred_in_",pred_in_d,"d_varimp_gbm_",fs_type,"_",pred_task,".rda"))
  
  #-------------------------------------------------------------------------------------------------------------
  lapse_tsk<-Sys.time()-start_tsk
  bm<-c(bm,paste0(lapse_tsk,units(lapse_tsk)))
  bm_nm<-c(bm_nm,"complete task")
  
  cat("\nFinish building reference models for task",pred_task,"in",lapse_tsk,units(lapse_tsk),".\n")
  
  #benchmark
  bm<-data.frame(bm_nm=bm_nm,bm_time=bm,
                 stringsAsFactors = F)
  saveRDS(bm,file=paste0("./data_local/model_ref/pred_in_",pred_in_d,"d_bm_gbm_",fs_type,"_",pred_task,".rda"))
  }
}

```


***

#### Objective 2.3: Benchmark Model Performance Evaluation
```{r}


```

***





