---
title: "Building and Validating Predictive Models for Acute Kidney Injury (AKI) using PCORnet CDM (Part II.1)"
author: "xing song"
date: "Feburary 09, 2019"
output: html_document
---
### Stage 2.2: Predictive Models Validation (Retrain)   

In this experiment, we will retrain the benchmark predictive model by quasi-replicating the model by [*Koyner et al*] for AKI risk prediction on the adult inpatients at each GPC site using PCORnet common data model. The model will be trained on 70% of the site's local data and validated on the remaining 30%.

[*Koyner et al*] https://www.ncbi.nlm.nih.gov/pubmed/29596073 

***

```{r setup, include=F}
#source utility functions
source("./R/util.R")
source("./R/var_etl_surv.R")

require_libraries(c("tidyr",
                    "dplyr",
                    "magrittr",
                    "stringr",
                    "broom",
                    "Matrix",
                    "xgboost",
                    "ROCR",
                    "PRROC",
                    "ResourceSelection",
                    "knitr",
                    "kableExtra",
                    "ggplot2",
                    "openxlsx"))

# experimental design parameters
#----prediction ending point
pred_end<-7

#-----prediction point
pred_in_d_opt<-c(1,2)

#-----prediction tasks
pred_task_lst<-c("stg1up","stg2up","stg3")

#-----feature selection type
fs_type_opt<-c("no_fs","rm_scr_bun")
rm_key<-c('2160-0','38483-4','14682-9','21232-4','35203-9','44784-7','59826-8',
          '16188-5','16189-3','59826-8','35591-7','50380-5','50381-3','35592-5',
          '44784-7','11041-1','51620-3','72271-0','11042-9','51619-5','35203-9','14682-9',
          '12966-8','12965-0','6299-2','59570-2','12964-3','49071-4','72270-2',
          '11065-0','3094-0','35234-4','14937-7',
          '3097-3','44734-2','BUN_SCR')
```


#### Preparation    

By running `Part I` of "render_report.R", the raw data tables should have already been collected and saved in the local `./data` folder (Note: these data tables are not visiable in the github ./data folder, but should be visible in the corresponding folder locally), that are

* `Table1.rda`: AKI patieht cohort table;       

* `AKI_DEMO.rda`: CDM demographic table cut for AKI cohort;   

* `AKI_VITAL.rda`: CDM vital table cut for AKI cohort;    

* `AKI_LAB.rda`: CDM lab table cut for AKI cohort;    

* `AKI_DX.rda`: CDM diagnosis table cut for AKI cohort;   

* `AKI_PX.rda`: CDM procedure table cut for AKI cohort;   

* `AKI_MED.rda`: CDM prescribing medication table cut for AKI cohort;   


***

#### Objective 2.1: Data Cleaning and Representation

In this section, the raw data tables will be cleaned and transformed to a discrete-survival-like representation, which will be used in the final modeling stage. To reduce the burden on memory requirments, the ETL (extract, transform, load) process will be performed in chunks with respect to **distinct prediction task, encounter years and variable types**. Meanwhile, indices for random paritioning will be assigned to each encounter. The ETL progress will be reported as follows:


```{r preprocess, include=F}

# collect and format variables on daily basis 
tbl1<-readRDS("./data//Table1.rda") %>%
  dplyr::mutate(yr=as.numeric(format(strptime(ADMIT_DATE, "%Y-%m-%d %H:%M:%S"),"%Y")))

for(pred_in_d in pred_in_d_opt){
  #--determine update time window
  tw<-as.double(seq(0,7))
  if(pred_in_d>1){
  tw<-tw[-seq_len(pred_in_d-1)]
  } 
  
  #--by chunks: encounter year
enc_yr<-tbl1 %>%
  dplyr::select(yr) %>%
  unique %>% arrange(yr) %>%
  filter(yr>2009) %>%
  unlist

#--by variable type
var_type<-c("demo","vital","lab","dx","px","med")

#--save results as array
for(pred_task in pred_task_lst){
  start_tsk<-Sys.time()
  cat("Start variable collection for task",pred_task,".\n")
  #---------------------------------------------------------------------------------------------
  
  var_by_yr<-list()
  var_bm<-list()
  rsample_idx<-c()
  for(i in seq_along(enc_yr)){
    start_i<-Sys.time()
    cat("...start variable collection for year",enc_yr[i],".\n")
    
    #--collect end_points
    dat_i<-tbl1 %>% filter(yr==enc_yr[i]) %>%
      dplyr::select(ENCOUNTERID,yr,
                    NONAKI_SINCE_ADMIT,
                    AKI1_SINCE_ADMIT,
                    AKI2_SINCE_ADMIT,
                    AKI3_SINCE_ADMIT) %>%
      gather(y,dsa_y,-ENCOUNTERID,-yr) %>%
      filter(!is.na(dsa_y)) %>%
      dplyr::mutate(y=recode(y,
                             "NONAKI_SINCE_ADMIT"=0,
                             "AKI1_SINCE_ADMIT"=1,
                             "AKI2_SINCE_ADMIT"=2,
                             "AKI3_SINCE_ADMIT"=3)) %>%
      dplyr::mutate(y=as.numeric(y))
    
    if(pred_task=="stg1up"){
      dat_i %<>%
        dplyr::mutate(y=as.numeric(y>0)) %>%
        group_by(ENCOUNTERID) %>% top_n(n=1L,wt=dsa_y) %>% ungroup
    }else if(pred_task=="stg2up"){
      dat_i %<>%
        # filter(y!=1) %>% # remove stage 1
        dplyr::mutate(y=as.numeric(y>1)) %>%
        group_by(ENCOUNTERID) %>% top_n(n=1L,wt=dsa_y) %>% ungroup
    }else if(pred_task=="stg3"){
      dat_i %<>%
        # filter(!(y %in% c(1,2))) %>% # remove stage 1,2
        dplyr::mutate(y=as.numeric(y>2)) %>%
        group_by(ENCOUNTERID) %>% top_n(n=1L,wt=dsa_y) %>% ungroup
    }else{
      stop("prediction task is not valid!")
    }
    
    #--random sampling
    rsample_idx %<>%
      bind_rows(dat_i %>% 
                  dplyr::select(ENCOUNTERID,yr) %>%
                  unique %>%
                  dplyr::mutate(cv10_idx=sample(1:10,n(),replace=T)))
    
    #--ETL variables
    X_surv<-c()
    y_surv<-c()
    var_etl_bm<-c()
    for(v in seq_along(var_type)){
      start_v<-Sys.time()
      
      #extract
      var_v<-readRDS(paste0("./data/AKI_",toupper(var_type[v]),".rda")) %>%
        semi_join(dat_i,by="ENCOUNTERID")
      
      if(var_type[v] != "demo"){
        if(var_type[v] == "med"){
          var_v %<>% 
            transform(value=strsplit(value,","),
                      dsa=strsplit(dsa,",")) %>%
            unnest(value,dsa) %>%
            dplyr::mutate(value=as.numeric(value),
                          dsa=as.numeric(dsa))
        }
        var_v %<>% filter(dsa <= pred_end)
      }
      
      #transform
      var_v<-format_data(dat=var_v,
                         type=var_type[v],
                         pred_end=pred_end)
      
      Xy_surv<-get_dsurv_temporal(dat=var_v,
                                  censor=dat_i,
                                  tw=tw,
                                  pred_in_d=pred_in_d)
      
      #load
      X_surv %<>% bind_rows(Xy_surv$X_surv) %>% unique
      y_surv %<>% bind_rows(Xy_surv$y_surv) %>% unique
      
      lapse_v<-Sys.time()-start_v
      var_etl_bm<-c(var_etl_bm,paste0(lapse_v,units(lapse_v)))
      cat("\n......finished ETL",var_type[v],"for year",enc_yr[i],"in",lapse_v,units(lapse_v),".\n")
    }
    var_by_yr[[i]]<-list(X_surv=X_surv,
                         y_surv=y_surv)
    
    lapse_i<-Sys.time()-start_i
    var_etl_bm<-c(var_etl_bm,paste0(lapse_i,units(lapse_i)))
    cat("\n...finished variabl collection for year",enc_yr[i],"in",lapse_i,units(lapse_i),".\n")
    
    var_bm[[i]]<-data.frame(bm_nm=c(var_type,"overall"),
                            bm_time=var_etl_bm,
                            stringsAsFactors = F)
  }
  #--save preprocessed data
  saveRDS(rsample_idx,file=paste0("./data/preproc/",pred_in_d,"d_rsample_idx_",pred_task,".rda"))
  saveRDS(var_by_yr,file=paste0("./data/preproc/",pred_in_d,"d_var_by_yr_",pred_task,".rda"))
  saveRDS(var_bm,file=paste0("./data/preproc/",pred_in_d,"d_var_bm",pred_task,".rda"))
  
  #---------------------------------------------------------------------------------------------
  lapse_tsk<-Sys.time()-start_tsk
  cat("\nFinish variable ETL for task:",pred_task,"in",pred_in_d,"days",",in",lapse_tsk,units(lapse_tsk),".\n")
  }
}
```

The final preprocessed intermediate tables from this code chunk should be found in the `./data/preproc/...` folder as the following intermediate data tables for different prediction tasks:   

* For AKI stage ≥ 1 in 24 hours: `1d_rsample_idx_stg1up.rda`, `1d_var_by_yr_stg1up.rda`, `1d_var_bm_stg1up.rda`   

* For AKI stage ≥ 2 in 24 hours: `1d_rsample_idx_stg2up.rda`, `1d_var_by_yr_stg2up.rda`, `1d_var_bm_stg2up.rda`   
 
* For AKI stage = 3 in 24 hours: `1d_rsample_idx_stg3.rda`, `1d_var_by_yr_stg3.rda`, `1d_var_bm_stg3.rda`   

* For AKI stage ≥ 1 in 48 hours: `2d_rsample_idx_stg1up.rda`, `2d_var_by_yr_stg1up.rda`, `1d_var_bm_stg1up.rda`   
 
* For AKI stage ≥ 2 in 48 hours: `2d_rsample_idx_stg2up.rda`, `2d_var_by_yr_stg2up.rda`, `1d_var_bm_stg2up.rda`   
 
* For AKI stage = 3 in 48 hours: `2d_rsample_idx_stg3.rda`, `2d_var_by_yr_stg3.rda`, `1d_var_bm_stg3.rda`   
 


***

#### Objective 2.2: Benchmark Model Development

We will adopt the AKI prediction model by [*Koyner et al*] using all variables from each site's CDM Demographic, Vital, Diagnosis, Procedure and Prescribing Medication tables.The same strategy as in Koyner et al for outlier removal and aggregation of repeated values have been followed. Training/Validation sets are partitioned based on pre-assigned indices in the files "..._rsample_idx_..." from previous part.  The model development progress will be reported as follows:


```{r benchmark, include=F}
for(pred_in_d in pred_in_d_opt){
  
  for(pred_task in pred_task_lst){
    bm<-c()
    bm_nm<-c()
    
    start_tsk<-Sys.time()
    # cat("Start build reference model for task",pred_task,".\n")
    #---------------------------------------------------------------------------------------------
    
    start_tsk_i<-Sys.time()
    #--prepare training and testing set
    yr_rg<-seq(2010,2018)
    X_tr<-c()
    X_ts<-c()
    y_tr<-c()
    y_ts<-c()
    rsample_idx<-readRDS(paste0("./data/preproc/",pred_in_d,"d_rsample_idx_",pred_task,".rda"))
    var_by_task<-readRDS(paste0("./data/preproc/",pred_in_d,"d_var_by_yr_",pred_task,".rda"))
    for(i in seq_along(yr_rg)){
      var_by_yr<-var_by_task[[i]]
      
      X_tr %<>% bind_rows(var_by_yr[["X_surv"]]) %>%
        semi_join(rsample_idx %>% filter(cv10_idx<=6 & yr<2017),
                  by="ENCOUNTERID")
      
      y_tr %<>% bind_rows(var_by_yr[["y_surv"]] %>%
                            left_join(rsample_idx %>% filter(cv10_idx<=6 & yr<2017),
                                      by="ENCOUNTERID"))
      
      X_ts %<>% bind_rows(var_by_yr[["X_surv"]]) %>%
        semi_join(rsample_idx %>% filter(cv10_idx>6 | yr>=2017),
                  by="ENCOUNTERID")
      
      y_ts %<>% bind_rows(var_by_yr[["y_surv"]] %>%
                            left_join(rsample_idx %>% filter(cv10_idx>6 | yr>=2017),
                                      by="ENCOUNTERID"))
      
      # cat("...finish stack data of encounters from",yr_rg[i],".\n")
    }
    lapse_i<-Sys.time()-start_tsk_i
    bm<-c(bm,paste0(lapse_i,units(lapse_i)))
    bm_nm<-c(bm_nm,"prepare data")
    
    for(fs_type in fs_type_opt){
        start_tsk_i<-Sys.time()
      #--pre-filter
      if(fs_type=="rm_scr_bun"){
        X_tr %<>%
          filter(!(key %in% c(rm_key,paste0(rm_key,"_change"))))
        
        X_ts %<>%
          filter(!(key %in% c(rm_key,paste0(rm_key,"_change"))))
      }
      
      #--transform training matrix
      y_tr %<>%
        filter(!is.na(cv10_idx)) %>%
        arrange(ENCOUNTERID,dsa_y) %>%
        unite("ROW_ID",c("ENCOUNTERID","dsa_y")) %>%
        arrange(ROW_ID) %>%
        unique
      
      X_tr %<>%
        arrange(ENCOUNTERID,dsa_y) %>%
        unite("ROW_ID",c("ENCOUNTERID","dsa_y")) %>%
        semi_join(y_tr,by="ROW_ID") %>%
        long_to_sparse_matrix(df=.,
                              id="ROW_ID",
                              variable="key",
                              val="value")
      
      #--collect variables used in training
      tr_key<-data.frame(key = unique(colnames(X_tr)),
                         stringsAsFactors = F)
      
      #--transform testing matrix
      y_ts %<>%
        filter(!is.na(cv10_idx)) %>%
        arrange(ENCOUNTERID,dsa_y) %>%
        unite("ROW_ID",c("ENCOUNTERID","dsa_y")) %>%
        arrange(ROW_ID) %>%
        unique
      
      X_ts %<>% 
        unite("ROW_ID",c("ENCOUNTERID","dsa_y")) %>%
        semi_join(y_ts,by="ROW_ID") %>%
        semi_join(tr_key,by="key")
      
      x_add<-tr_key %>%
        anti_join(data.frame(key = unique(X_ts$key),
                             stringsAsFactors = F),
                  by="key")
      
      #align with training
      if(nrow(x_add)>0){
        X_ts %<>%
          arrange(ROW_ID) %>%
          bind_rows(data.frame(ROW_ID = rep("0_0",nrow(x_add)),
                               dsa = -99,
                               key = x_add$key,
                               value = 0,
                               stringsAsFactors=F))
      }
      X_ts %<>%
        long_to_sparse_matrix(df=.,
                              id="ROW_ID",
                              variable="key",
                              val="value")
      if(nrow(x_add)>0){
        X_ts<-X_ts[-1,]
      }
      
      #check alignment
      if(!all(row.names(X_tr)==y_tr$ROW_ID)){
        stop("row ids of traning set don't match!")
      }
      if(!all(row.names(X_ts)==y_ts$ROW_ID)){
        stop("row ids of testing set don't match!")
      }
      if(!all(colnames(X_tr)==colnames(X_ts))){
        stop("feature names don't match!")
      }
      
      #--covert to xgb data frame
      dtrain<-xgb.DMatrix(data=X_tr,label=y_tr$y)
      dtest<-xgb.DMatrix(data=X_ts,label=y_ts$y)
      
      lapse_i<-Sys.time()-start_tsk_i
      bm<-c(bm,paste0(lapse_i,units(lapse_i)))
      bm_nm<-c(bm_nm,"transform data")
      
      # cat("...finish formatting training and testing sets.\n")
      
      start_tsk_i<-Sys.time()
      #--get indices for k folds
      y_tr %<>% dplyr::mutate(row_idx = 1:n())
      folds<-list()
      for(fd in seq_len(max(y_tr$cv10_idx))){
        fd_df<-y_tr %>% 
          filter(cv10_idx==fd) %>%
          dplyr::select(row_idx)
        folds[[fd]]<-fd_df$row_idx
      }
      
      #--tune hyperparameter
      #hyper-parameter grid for xgboost
      eval_metric<-"auc"
      objective<-"binary:logistic"
      grid_params<-expand.grid(
        max_depth=c(4,10),
        eta=c(0.1,0.01),
        min_child_weight=1,
        subsample=0.8,
        colsample_bytree=0.8, 
        gamma=c(1,10)
      )
      
      verb<-TRUE
      bst_grid<-c()
      bst_grid_cv<-c()
      metric_name<-paste0("test_", eval_metric,"_mean")
      metric_sd_name<-paste0("test_", eval_metric,"_std")
      
      for(i in seq_len(dim(grid_params)[1])){
        start_i<-Sys.time()
        param<-as.list(grid_params[i,])
        # param$scale_pos_weight=mean(train$y_train$DKD_IND_additive) #inbalance sampling
        param$scale_pos_weight=1 #balance sampling
        
        bst <- xgb.cv(param,
                      dtrain,
                      objective = objective,
                      metrics = eval_metric,
                      maximize = TRUE,
                      nrounds=2000,
                      # nfold = 5,
                      folds = folds,
                      early_stopping_rounds = 100,
                      print_every_n = 100,
                      prediction = T) #keep cv results
        
        bst_grid<-rbind(bst_grid, cbind(grid_params[i,],
                                        metric=max(bst$evaluation_log[[metric_name]]),
                                        steps=which(bst$evaluation_log[[metric_name]]==max(bst$evaluation_log[[metric_name]]))[1]))
        
        bst_grid_cv<-cbind(bst_grid_cv,bst$pred)
        
        if(verb){
          cat('...finished train case:',paste0(paste0(c(colnames(grid_params),"scale_pos_weight"),"="),param,collapse="; "),
              'in',Sys.time()-start_i,units(Sys.time()-start_i),"\n")
          start_i<-Sys.time()
        }
      }
      hyper_param<-bst_grid[which.max(bst_grid$metric),]
      
      lapse_i<-Sys.time()-start_tsk_i
      bm<-c(bm,paste0(lapse_i,units(lapse_i)))
      bm_nm<-c(bm_nm,"tune model")
      
      # cat("...finish model tunning.\n")
      
      start_tsk_i<-Sys.time()  
      #--validation
      xgb_tune<-xgb.train(data=dtrain,
                          max_depth=hyper_param$max_depth,
                          maximize = TRUE,
                          eta=hyper_param$eta,
                          nrounds=hyper_param$steps,
                          eval_metric="auc",
                          objective="binary:logistic",
                          print_every_n = 100)
      
      valid<-data.frame(y_ts,
                        pred = predict(xgb_tune,dtest),
                        stringsAsFactors = F)
      
      #--feature importance
      feat_imp<-xgb.importance(colnames(X_tr),model=xgb_tune)
      
      lapse_i<-Sys.time()-start_tsk_i
      bm<-c(bm,paste0(lapse_i,units(lapse_i)))
      bm_nm<-c(bm_nm,"validate model")
      
      # cat("...finish model validating.\n")
      
      #--save model and other results
      saveRDS(xgb_tune,file=paste0("./data/model_ref/pred_in_",pred_in_d,"d_model_gbm_",fs_type,"_",pred_task,".rda"))
      saveRDS(bst_grid,file=paste0("./data/model_ref/pred_in_",pred_in_d,"d_hyperpar_gbm_",fs_type,"_",pred_task,".rda"))
      saveRDS(valid,file=paste0("./data/model_ref/pred_in_",pred_in_d,"d_valid_gbm_",fs_type,"_",pred_task,".rda"))
      saveRDS(feat_imp,file=paste0("./data/model_ref/pred_in_",pred_in_d,"d_varimp_gbm_",fs_type,"_",pred_task,".rda"))
      
      #-------------------------------------------------------------------------------------------------------------
      lapse_tsk<-Sys.time()-start_tsk
      bm<-c(bm,paste0(lapse_tsk,units(lapse_tsk)))
      bm_nm<-c(bm_nm,"complete task")
      
      cat("\nFinish building reference models for task:",pred_task,"in",pred_in_d,"with",fs_type,",in",lapse_tsk,units(lapse_tsk),".\n")
      
      #benchmark
      bm<-data.frame(bm_nm=bm_nm,bm_time=bm,
                     stringsAsFactors = F)
      saveRDS(bm,file=paste0("./data/model_ref/pred_in_",pred_in_d,"d_bm_gbm_",fs_type,"_",pred_task,".rda"))
    }
  }
}

```

For each prediction task, defined as "predict AKI stage X in Y days, with/without Scr", 4 intermedicate data files have been generated and saved in `./data/model_ref/...`, which are:  

* `..._hyperpar_gbm_...rda`: the final hyper-parameter sets after tunning;   

* `..._model_gbm_...rda`: the final gbm model after tunning;   

* `..._valid_gbm_...rda`: the predictted probability on validataion set;      
 
* `..._varimp_gbm_...rda`: the final list of variable importance     


***


#### Objective 2.3: Performance Evaluations for Benchmark Model

```{r perf_summ, include=F}
rm(list=c("X_ts","y_ts","dtest","X_tr","y_tr","dtrain")); gc() #release some memory

for(pred_in_d in pred_in_d_opt){
  for(fs_type in fs_type_opt){
    perf_tbl_full<-c()
    perf_tbl<-c()
    calib_tbl<-c()
    varimp_tbl<-c()
    for(i in seq_along(pred_task_lst)){
      valid<-readRDS(paste0("./data/model_ref/pred_in_",pred_in_d,"d_valid_gbm_",fs_type,"_",pred_task_lst[i],".rda"))
      
      #overall summary
      perf_summ<-get_perf_summ(pred=valid$pred,
                               real=valid$y,
                               keep_all_cutoffs=T)
      perf_tbl_full %<>% 
        bind_rows(perf_summ$perf_at %>% 
                    dplyr::mutate(pred_task=pred_task_lst[i],pred_in_d=pred_in_d,fs_type=fs_type))
      
      perf_tbl %<>% 
        bind_rows(perf_summ$perf_summ %>% 
                    dplyr::mutate(pred_task=pred_task_lst[i],pred_in_d=pred_in_d,fs_type=fs_type))
      
      #calibration
      calib<-get_calibr(pred=valid$pred,
                       real=valid$y,
                       n_bin=20)
      
      calib_tbl %<>% 
        bind_rows(calib %>% 
                    dplyr::mutate(pred_task=pred_task_lst[i],pred_in_d=pred_in_d,fs_type=fs_type))
      
      #variable
      varimp<-readRDS(paste0("./data/model_ref/pred_in_",pred_in_d,"d_varimp_gbm_",fs_type,"_",pred_task_lst[i],".rda")) %>%
        dplyr::mutate(rank=1:n(),
                      Gain_rescale=round(Gain/Gain[1]*100)) %>%
        dplyr::select(rank,Feature,Gain_rescale)
      
      varimp_tbl %<>% 
        bind_rows(varimp %>% 
                    mutate(pred_task=pred_task_lst[i],pred_in_d=pred_in_d,fs_type=fs_type,tot_feature=nrow(varimp)))
    }
    
    perf_out<-list(perf_tbl_full=perf_tbl_full,
                   perf_tbl=perf_tbl,
                   calib_tbl=calib_tbl,
                   varimp_tbl=varimp_tbl)
    
    #save results as r data.frame
    saveRDS(perf_out,file=paste0("./data/model_ref/pred_in_",pred_in_d,"d_",fs_type,"_baseline_model_perf.rda"))
  }
}

```


```{r data_dict, include=F}
# plot variable importance
data_dict<-read.csv("./ref/feature_dict.csv") %>%
  dplyr::select(FIELD_NAME,VALUESET_ITEM,VALUESET_ITEM_DESCRIPTOR) %>%
  mutate(VALUESET_ITEM_DESCRIPTOR=as.character(VALUESET_ITEM_DESCRIPTOR)) %>%
  mutate(VALUESET_ITEM=gsub("'","",VALUESET_ITEM))

tbl1<-readRDS("./data/Table1.rda") %>%
  dplyr::select(ENCOUNTERID,SERUM_CREAT_BASE) %>%
  inner_join(readRDS("./data/AKI_DEMO.rda") %>%
              filter(key=="AGE") %>%
              dplyr::select(ENCOUNTERID,key,value) %>%
              spread(key,value) %>%
              mutate(AGE=as.numeric(AGE)),
            by="ENCOUNTERID")

final_out<-list()
```


#### Task I: 24-hour advanced prediction

```{r pred24hr, include=F}
pred_in_d<-1
```


*(I.a) Overall Prediction Performance and Subgroup Analysis*   

```{r pred24_summ,echo=F}
fs_type_opt<-c("no_fs","rm_scr_bun")

#overall summary
perf_overall<-c()
for(fs_type in fs_type_opt){
  perf_overall %<>%
    bind_rows(readRDS(paste0("./data/model_ref/pred_in_",pred_in_d,"d_",fs_type,"_baseline_model_perf.rda"))$perf_tbl %>%
                filter(overall_meas %in% c("roauc",
                                           "roauc_low",
                                           "roauc_up",
                                           "prauc1",
                                           "opt_sens",
                                           "opt_spec",
                                           "opt_ppv",
                                           "opt_npv")) %>%
                dplyr::mutate(overall_meas=recode(overall_meas,
                                                  roauc="1.ROAUC",
                                                  prauc1="2.PRAUC",
                                                  opt_sens="3.Optimal Sensitivity",
                                                  opt_spec="4.Optimal Specificity",
                                                  opt_ppv="5.Optimal Positive Predictive Value",
                                                  opt_npv="6.Optimal Negative Predictive Value"),
                              grp="Overall",
                              subgrp=fs_type))
}

#subgroup-age
fs_type<-"no_fs"
subgrp_age<-c()
for(i in seq_along(pred_task_lst)){
      valid<-readRDS(paste0("./data/model_ref/pred_in_",pred_in_d,"d_valid_gbm_",fs_type,"_",pred_task_lst[i],".rda")) %>%
        mutate(ENCOUNTERID=gsub("_.*","",ROW_ID),
               day_at=as.numeric(gsub(".*_","",ROW_ID))) %>%
        inner_join(tbl1,by="ENCOUNTERID") %>%
        dplyr::select(ENCOUNTERID,day_at,y,pred,AGE) %>%
        dplyr::mutate(age=cut(AGE,breaks=c(0,45,65,Inf),include.lowest=T,right=F))
      
      grp_vec<-unique(valid$age)
      for(grp in seq_along(grp_vec)){
        valid_grp<-valid %>% filter(age==grp_vec[grp])
        grp_summ<-get_perf_summ(pred=valid_grp$pred,
                                real=valid_grp$y,
                                keep_all_cutoffs=F)$perf_summ %>%
          dplyr::filter(overall_meas %in% c("roauc",
                                            "roauc_low",
                                            "roauc_up",
                                            "prauc1",
                                            "opt_sens",
                                            "opt_spec",
                                            "opt_ppv",
                                            "opt_npv")) %>%
          dplyr::mutate(overall_meas=recode(overall_meas,
                                            roauc="1.ROAUC",
                                            prauc1="2.PRAUC",
                                            opt_sens="3.Optimal Sensitivity",
                                            opt_spec="4.Optimal Specificity",
                                            opt_ppv="5.Optimal Positive Predictive Value",
                                            opt_npv="6.Optimal Negative Predictive Value"))
        subgrp_age %<>% 
          bind_rows(grp_summ %>% 
                      dplyr::mutate(grp="Subgrp_AGE",
                                    subgrp=grp_vec[grp],
                                    pred_task=paste0("stg",i,"up")))
      }
  }
  
  #subgroup-scr
  subgrp_scr<-c()
  for(i in seq_along(pred_task_lst)){
      valid<-readRDS(paste0("./data/model_ref/pred_in_",pred_in_d,"d_valid_gbm_",fs_type,"_",pred_task_lst[i],".rda")) %>%
        mutate(ENCOUNTERID=gsub("_.*","",ROW_ID),
               day_at=as.numeric(gsub(".*_","",ROW_ID))) %>%
        inner_join(tbl1,by="ENCOUNTERID") %>%
        dplyr::select(ENCOUNTERID,day_at,y,pred,SERUM_CREAT_BASE) %>%
        mutate(admit_scr=cut(SERUM_CREAT_BASE,breaks=c(0,1,2,3,Inf),include.lowest=T,right=F))
      
      grp_vec<-unique(valid$admit_scr)
      for(grp in seq_along(grp_vec)){
        valid_grp<-valid %>% filter(admit_scr==grp_vec[grp])
        grp_summ<-get_perf_summ(pred=valid_grp$pred,
                                real=valid_grp$y,
                                keep_all_cutoffs=F)$perf_summ %>%
          filter(overall_meas %in% c("roauc",
                                     "roauc_low",
                                     "roauc_up",
                                     "prauc1",
                                     "opt_sens",
                                     "opt_spec",
                                     "opt_ppv",
                                     "opt_npv")) %>%
          dplyr::mutate(overall_meas=recode(overall_meas,
                                            roauc="1.ROAUC",
                                            prauc1="2.PRAUC",
                                            opt_sens="3.Optimal Sensitivity",
                                            opt_spec="4.Optimal Specificity",
                                            opt_ppv="5.Optimal Positive Predictive Value",
                                            opt_npv="6.Optimal Negative Predictive Value"))
        subgrp_scr %<>% 
          bind_rows(grp_summ %>% 
                      dplyr::mutate(grp="Subgrp_Scr_Base",
                                    subgrp=grp_vec[grp],
                                    pred_task=paste0("stg",i,"up")))
      }
  }

perf_overall %<>%
  bind_rows(subgrp_age %>% 
              dplyr::mutate(pred_in_d=pred_in_d,fs_type=fs_type)) %>%
  bind_rows(subgrp_scr %>% 
              dplyr::mutate(pred_in_d=pred_in_d,fs_type=fs_type)) %>%
  dplyr::select(pred_in_d,fs_type,grp,subgrp,pred_task,overall_meas,meas_val) %>%
  dplyr::mutate(pred_task=recode(pred_task,
                                 `stg1up`="a.AKI>=1",
                                 `stg2up`="b.AKI>=2",
                                 `stg3`="c.AKI=3",
                                 `stg3up`="c.AKI=3"),
                meas_val=round(meas_val,4)) %>%
                spread(overall_meas,meas_val) %>%
                dplyr::mutate(roauc_up=paste0(roauc_up,")")) %>%
                unite("1.ROAUC_CI",c("roauc_low","roauc_up"),sep=",") %>%
                unite("1.ROAUC",c("1.ROAUC","1.ROAUC_CI"),sep=" (")  

row_grp_pos<-perf_overall %>% 
  arrange(grp,pred_task) %>%
  mutate(rn=1:n()) %>%
  dplyr::mutate(root_grp=gsub(":.*","",grp)) %>%
  group_by(root_grp,pred_in_d) %>%
  dplyr::summarize(begin=rn[1],
                   end=rn[n()]) %>%
  ungroup

kable(perf_overall %>% arrange(grp,pred_task),
      caption="Table1 - 24-Hour Prediction of AKI1, AKI2, AKI3") %>%
  kable_styling("striped", full_width = F) %>%
  group_rows("Overall", row_grp_pos$begin[1],row_grp_pos$end[1]) %>%
  group_rows("Subgroup-Age", row_grp_pos$begin[2],row_grp_pos$end[2]) %>%
  group_rows("Subgroup-Scr_Base", row_grp_pos$begin[3],row_grp_pos$end[3])

final_out[["pred24_summ"]]<-perf_overall
```


*(I.b) Full metric paths at various cutoff points*

```{r pred24_cutoff, echo=F,fig.height = 5, fig.width = 12}
#plot out sens, spec, ppv, npv on a scale of cutoff probabilities
brks<-unique(c(0,seq(0.001,0.01,by=0.001),seq(0.02,0.1,by=0.01),seq(0.2,1,by=0.1)))
pred24_cutoff_plot<-list()

for(fs_type_i in seq_along(fs_type_opt)){
  fs_type<-fs_type_opt[fs_type_i]
  
  perf_cutoff<-readRDS(paste0("./data/model_ref/pred_in_",pred_in_d,"d_",fs_type,"_baseline_model_perf.rda"))$perf_tbl_full %>%
    dplyr::select(cutoff,rec_sens,spec,ppv,npv,pred_task) %>%
    mutate(bin=cut(cutoff,breaks=brks,include.lowest=T,label=F)) %>%
    group_by(bin,pred_task) %>%
    dplyr::summarise(cutoff=round(max(cutoff),3),
                     sens_l=quantile(rec_sens,0.025,na.rm=T),
                     sens=median(rec_sens,na.rm=T),
                     sens_u=quantile(rec_sens,0.975,na.rm=T),
                     spec_l=quantile(spec,0.025,na.rm=T),
                     spec=median(spec,na.rm=T),
                     spec_u=quantile(spec,0.975,na.rm=T),
                     ppv_l=quantile(ppv,0.025,na.rm=T),
                     ppv=median(ppv,na.rm=T),
                     ppv_u=quantile(ppv,0.975,na.rm=T),
                     npv_l=quantile(npv,0.025,na.rm=T),
                     npv=median(npv,na.rm=T),
                     npv_u=quantile(npv,0.975,na.rm=T)) %>%
    ungroup %>% dplyr::select(-bin) %>%
    gather(metric_type,metric_val,-cutoff,-pred_task) %>%
    mutate(metric_type2=ifelse(grepl("_l",metric_type),"low",
                               ifelse(grepl("_u",metric_type),"up","mid")),
          metric_type=gsub("_.*","",metric_type)) %>%
    spread(metric_type2,metric_val) %>%
    mutate(pred_task=recode(pred_task,
                            `stg1up`="a.AKI>=1",
                            `stg2up`="b.AKI>=2",
                            `stg3`="c.AKI=3"))
  
  pred24_cutoff_plot[[fs_type_i]]<-ggplot(perf_cutoff %>% dplyr::filter(cutoff <=0.15),
                   aes(x=cutoff,y=mid,color=metric_type,fill=metric_type))+
    geom_line()+ geom_ribbon(aes(ymin=low,ymax=up),alpha=0.3)+
    labs(x="cutoff probability",y="performance metrics",
         title=paste0("Figure",fs_type_i," - Metrics at different cutoff points:",fs_type))+
    facet_wrap(~pred_task,scales="free",ncol=3)
  
  final_out[[paste0("pred24_cutoff_",fs_type)]]<-perf_cutoff
}

print(pred24_cutoff_plot[[1]])
print(pred24_cutoff_plot[[2]])
```


*(I.c) Calibration Plots*

```{r pred24_calib, echo=F, fig.height = 5, fig.width = 12}
pred24_calib_plot<-list()

# plot calibration
for(fs_type_i in seq_along(fs_type_opt)){
  fs_type<-fs_type_opt[fs_type_i]
  
  calib_tbl<-readRDS(paste0("./data/model_ref/pred_in_",pred_in_d,"d_",fs_type,"_baseline_model_perf.rda"))$calib_tbl %>%
    mutate(pred_task=recode(pred_task,
                          `stg1up`="a.AKI>=1",
                          `stg2up`="b.AKI>=2",
                          `stg3`="c.AKI=3"))
  
  pred24_calib_plot[[fs_type_i]]<-ggplot(calib_tbl,aes(x=y_p,y=pred_p))+
    geom_point()+geom_abline(intercept=0,slope=1)+
    geom_errorbar(aes(ymin=binCI_lower,ymax=binCI_upper))+
    labs(x="Actual Probability",y="Predicted Probability",
         title=paste0("Figure",fs_type_i+2," - Calibration:",fs_type))+
    facet_wrap(~pred_task,scales="free")
  
  final_out[[paste0("pred24_calibr_",fs_type)]]<-calib_tbl
}
print(pred24_calib_plot[[1]])
print(pred24_calib_plot[[2]])
```


*(I.d) Top Importance Variables*

```{r pred24_varimp, echo=F,fig.height = 12, fig.width = 8}
pred24_varimp_plot<-list()

# plot variable importance
for(fs_type_i in seq_along(fs_type_opt)){
  fs_type<-fs_type_opt[fs_type_i]
  
  varimp<-readRDS(paste0("./data/model_ref/pred_in_",pred_in_d,"d_",fs_type,"_baseline_model_perf.rda"))$varimp_tbl %>%
  dplyr::mutate(suffix=gsub("_","",str_extract(Feature,"((\\_min)|(\\_slope)|(\\_change)|(\\_cum))+"))) %>%
  dplyr::mutate(feat=gsub("((\\_min)|(\\_slope)|(\\_change)|(\\_cum))+","",Feature)) %>%
  left_join(data_dict,by=c("feat"="VALUESET_ITEM")) %>%
  dplyr::mutate(feat = ifelse(is.na(VALUESET_ITEM_DESCRIPTOR),Feature,VALUESET_ITEM_DESCRIPTOR)) %>%
  dplyr::mutate(feat = ifelse(!is.na(suffix),paste0(feat,"_",suffix),feat)) %>%
  dplyr::mutate(feat_rank=paste0(ifelse(rank<10,
                                        paste0("0",rank),
                                        as.character(rank)),
                                 ".",feat)) %>%
  unique %>%
  mutate(feat_rank=paste0(substr(feat_rank,1,50),"...")) %>%
  dplyr::mutate(feat_rank=as.factor(feat_rank)) %>%
  dplyr::mutate(feat_rank=factor(feat_rank,levels=rev(levels(feat_rank)))) %>%
  mutate(pred_task=recode(pred_task,
                          `stg1up`="a.AKI>=1",
                          `stg2up`="b.AKI>=2",
                          `stg3`="c.AKI=3"))

  pred24_varimp_plot[[fs_type_i]]<-ggplot(varimp %>% filter(rank <= 20),
                                          aes(x=feat_rank,y=Gain_rescale))+
    geom_bar(stat="identity")+
    labs(x="Features",y="Normalized Scale",
         title=paste0("Figure",fs_type_i+4," - Top Important Variables:",fs_type))+
    coord_flip()+scale_y_continuous(trans = "reverse")+
    facet_wrap(~pred_task,scales = "free",ncol=1)
  
  final_out[[paste0("pred24_varimp_",fs_type)]]<-varimp
}
print(pred24_varimp_plot[[1]])
print(pred24_varimp_plot[[2]])
```


***

#### Task II: 48-hour advanced prediction

```{r pred48hr, include=F}
pred_in_d<-2
```

*(II.a) Overall Prediction Performance*

```{r pred48_summ,echo=F}
fs_type_opt<-c("no_fs","rm_scr_bun")

#overall summary
perf_overall<-c()
for(fs_type in fs_type_opt){
  perf_overall %<>%
    bind_rows(readRDS(paste0("./data/model_ref/pred_in_",pred_in_d,"d_",fs_type,"_baseline_model_perf.rda"))$perf_tbl %>%
                filter(overall_meas %in% c("roauc",
                                           "roauc_low",
                                           "roauc_up",
                                           "prauc1",
                                           "opt_sens",
                                           "opt_spec",
                                           "opt_ppv",
                                           "opt_npv")) %>%
                dplyr::mutate(overall_meas=recode(overall_meas,
                                                  roauc="1.ROAUC",
                                                  prauc1="2.PRAUC",
                                                  opt_sens="3.Optimal Sensitivity",
                                                  opt_spec="4.Optimal Specificity",
                                                  opt_ppv="5.Optimal Positive Predictive Value",
                                                  opt_npv="6.Optimal Negative Predictive Value"),
                              grp="Overall",
                              subgrp=fs_type))
}

#subgroup-age
fs_type<-"no_fs"
subgrp_age<-c()
for(i in seq_along(pred_task_lst)){
      valid<-readRDS(paste0("./data/model_ref/pred_in_",pred_in_d,"d_valid_gbm_",fs_type,"_",pred_task_lst[i],".rda")) %>%
        mutate(ENCOUNTERID=gsub("_.*","",ROW_ID),
               day_at=as.numeric(gsub(".*_","",ROW_ID))) %>%
        inner_join(tbl1,by="ENCOUNTERID") %>%
        dplyr::select(ENCOUNTERID,day_at,y,pred,AGE) %>%
        dplyr::mutate(age=cut(AGE,breaks=c(0,45,65,Inf),include.lowest=T,right=F))
      
      grp_vec<-unique(valid$age)
      for(grp in seq_along(grp_vec)){
        valid_grp<-valid %>% filter(age==grp_vec[grp])
        grp_summ<-get_perf_summ(pred=valid_grp$pred,
                                real=valid_grp$y,
                                keep_all_cutoffs=F)$perf_summ %>%
          dplyr::filter(overall_meas %in% c("roauc",
                                            "roauc_low",
                                            "roauc_up",
                                            "prauc1",
                                            "opt_sens",
                                            "opt_spec",
                                            "opt_ppv",
                                            "opt_npv")) %>%
          dplyr::mutate(overall_meas=recode(overall_meas,
                                            roauc="1.ROAUC",
                                            prauc1="2.PRAUC",
                                            opt_sens="3.Optimal Sensitivity",
                                            opt_spec="4.Optimal Specificity",
                                            opt_ppv="5.Optimal Positive Predictive Value",
                                            opt_npv="6.Optimal Negative Predictive Value"))
        subgrp_age %<>% 
          bind_rows(grp_summ %>% 
                      dplyr::mutate(grp="Subgrp_AGE",
                                    subgrp=grp_vec[grp],
                                    pred_task=paste0("stg",i,"up")))
      }
  }
  
  #subgroup-scr
  subgrp_scr<-c()
  for(i in seq_along(pred_task_lst)){
      valid<-readRDS(paste0("./data/model_ref/pred_in_",pred_in_d,"d_valid_gbm_",fs_type,"_",pred_task_lst[i],".rda")) %>%
        mutate(ENCOUNTERID=gsub("_.*","",ROW_ID),
               day_at=as.numeric(gsub(".*_","",ROW_ID))) %>%
        inner_join(tbl1,by="ENCOUNTERID") %>%
        dplyr::select(ENCOUNTERID,day_at,y,pred,SERUM_CREAT_BASE) %>%
        mutate(admit_scr=cut(SERUM_CREAT_BASE,breaks=c(0,1,2,3,Inf),include.lowest=T,right=F))
      
      grp_vec<-unique(valid$admit_scr)
      for(grp in seq_along(grp_vec)){
        valid_grp<-valid %>% filter(admit_scr==grp_vec[grp])
        grp_summ<-get_perf_summ(pred=valid_grp$pred,
                                real=valid_grp$y,
                                keep_all_cutoffs=F)$perf_summ %>%
          filter(overall_meas %in% c("roauc",
                                     "roauc_low",
                                     "roauc_up",
                                     "prauc1",
                                     "opt_sens",
                                     "opt_spec",
                                     "opt_ppv",
                                     "opt_npv")) %>%
          dplyr::mutate(overall_meas=recode(overall_meas,
                                            roauc="1.ROAUC",
                                            prauc1="2.PRAUC",
                                            opt_sens="3.Optimal Sensitivity",
                                            opt_spec="4.Optimal Specificity",
                                            opt_ppv="5.Optimal Positive Predictive Value",
                                            opt_npv="6.Optimal Negative Predictive Value"))
        subgrp_scr %<>% 
          bind_rows(grp_summ %>% 
                      dplyr::mutate(grp="Subgrp_Scr_Base",
                                    subgrp=grp_vec[grp],
                                    pred_task=paste0("stg",i,"up")))
      }
  }

perf_overall %<>%
  bind_rows(subgrp_age %>% 
              dplyr::mutate(pred_in_d=pred_in_d,fs_type=fs_type)) %>%
  bind_rows(subgrp_scr %>% 
              dplyr::mutate(pred_in_d=pred_in_d,fs_type=fs_type)) %>%
  dplyr::select(pred_in_d,fs_type,grp,pred_task,overall_meas,meas_val) %>%
  dplyr::mutate(pred_task=recode(pred_task,
                                 `stg1up`="a.AKI>=1",
                                 `stg2up`="b.AKI>=2",
                                 `stg3`="c.AKI=3",
                                 `stg3up`="c.AKI=3"),
                meas_val=round(meas_val,4)) %>%
                spread(overall_meas,meas_val) %>%
                dplyr::mutate(roauc_up=paste0(roauc_up,")")) %>%
                unite("1.ROAUC_CI",c("roauc_low","roauc_up"),sep=",") %>%
                unite("1.ROAUC",c("1.ROAUC","1.ROAUC_CI"),sep=" (")  

row_grp_pos<-perf_overall %>% 
  arrange(grp,pred_task) %>%
  mutate(rn=1:n()) %>%
  dplyr::mutate(root_grp=gsub(":.*","",grp)) %>%
  group_by(root_grp,pred_in_d) %>%
  dplyr::summarize(begin=rn[1],
                   end=rn[n()]) %>%
  ungroup

kable(perf_overall %>% arrange(grp,pred_task),
      caption="Table1 - 48-Hour Prediction of AKI1, AKI2, AKI3") %>%
  kable_styling("striped", full_width = F) %>%
  group_rows("Overall", row_grp_pos$begin[1],row_grp_pos$end[1]) %>%
  group_rows("Subgroup-Age", row_grp_pos$begin[2],row_grp_pos$end[2]) %>%
  group_rows("Subgroup-Scr_Base", row_grp_pos$begin[3],row_grp_pos$end[3])

final_out[["pred48_summ"]]<-perf_overall
```


*(II.b) Full metric paths at various cutoff points*

```{r pred48_cutoff,echo=F, fig.height = 5, fig.width = 12}
pred48_cutoff_plot<-list()

#plot out sens, spec, ppv, npv on a scale of cutoff probabilities
brks<-unique(c(0,seq(0.001,0.01,by=0.001),seq(0.02,0.1,by=0.01),seq(0.2,1,by=0.1)))
for(fs_type_i in seq_along(fs_type_opt)){
  fs_type<-fs_type_opt[fs_type_i]
  
  perf_cutoff<-readRDS(paste0("./data/model_ref/pred_in_",pred_in_d,"d_",fs_type,"_baseline_model_perf.rda"))$perf_tbl_full %>%
  dplyr::select(cutoff,size,rec_sens,spec,ppv,npv,pred_task) %>%
  mutate(bin=cut(cutoff,breaks=brks,include.lowest=T,label=F)) %>%
  group_by(bin,pred_task) %>%
  dplyr::summarise(size=sum(size),
                   cutoff=round(max(cutoff),3),
                   sens_l=quantile(rec_sens,0.025,na.rm=T),
                   sens=median(rec_sens,na.rm=T),
                   sens_u=quantile(rec_sens,0.975,na.rm=T),
                   spec_l=quantile(spec,0.025,na.rm=T),
                   spec=median(spec,na.rm=T),
                   spec_u=quantile(spec,0.975,na.rm=T),
                   ppv_l=quantile(ppv,0.025,na.rm=T),
                   ppv=median(ppv,na.rm=T),
                   ppv_u=quantile(ppv,0.975,na.rm=T),
                   npv_l=quantile(npv,0.025,na.rm=T),
                   npv=median(npv,na.rm=T),
                   npv_u=quantile(npv,0.975,na.rm=T)) %>%
  ungroup %>%
  group_by(pred_task) %>%
  dplyr::mutate(size_overall=sum(size),
                size_cum=cumsum(size)) %>%
  ungroup %>%
  mutate(size_abv=size_overall-size_cum) %>%
  dplyr::select(-bin,-size_overall,-size_cum,-size) %>%
  gather(metric_type,metric_val,-cutoff,-size_abv,-pred_task) %>%
  mutate(metric_type2=ifelse(grepl("_l",metric_type),"low",
                             ifelse(grepl("_u",metric_type),"up","mid")),
         metric_type=gsub("_.*","",metric_type)) %>%
  spread(metric_type2,metric_val) %>%
  mutate(pred_task=recode(pred_task,
                          `stg1up`="a.At least AKI1",
                          `stg2up`="b.At least AKI2",
                          `stg3`="c.AKI3"))

  pred48_cutoff_plot[[fs_type_i]]<-ggplot(perf_cutoff %>% dplyr::filter(cutoff <=0.15),
                                          aes(x=cutoff,y=mid,color=metric_type,fill=metric_type))+
    geom_line()+ geom_ribbon(aes(ymin=low,ymax=up),alpha=0.3)+
    labs(x="cutoff probability",y="performance metrics",
         title=paste0("Figure",fs_type_i+6," - Metrics at different cutoff points:",fs_type))+
    facet_wrap(~pred_task,scales="free",ncol=3)
  
  final_out[[paste0("pred48_cutoff_",fs_type)]]<-perf_cutoff
}
print(pred48_cutoff_plot[[1]])
print(pred48_cutoff_plot[[2]])
```


*(II.c) Calibration Plots*

```{r pred48_calib,echo=F, fig.height = 5, fig.width = 12}
pred48_calib_plot<-list()

# plot calibration
for(fs_type_i in seq_along(fs_type_opt)){
  fs_type<-fs_type_opt[fs_type_i]
  
  calib_tbl<-readRDS(paste0("./data/model_ref/pred_in_",pred_in_d,"d_",fs_type,"_baseline_model_perf.rda"))$calib_tbl %>%
  mutate(pred_task=recode(pred_task,
                          `stg1up`="a.AKI>=1",
                          `stg2up`="b.AKI>=2",
                          `stg3`="c.AKI=3"))
  
  pred48_calib_plot[[fs_type_i]]<-ggplot(calib_tbl,aes(x=y_p,y=pred_p))+
    geom_point()+geom_abline(intercept=0,slope=1)+
    geom_errorbar(aes(ymin=binCI_lower,ymax=binCI_upper))+
    labs(x="Actual Probability",y="Predicted Probability",
         title=paste0("Figure",fs_type_i+8," - Calibration:",fs_type))+
    facet_wrap(~pred_task,scales="free")
  
  final_out[[paste0("pred48_calibr_",fs_type)]]<-calib_tbl
}
print(pred48_calib_plot[[1]])
print(pred48_calib_plot[[2]])
```


*(II.d) Top Importance Variables*

```{r pred48_varimp, echo=F, fig.height = 12, fig.width = 8}
pred48_varimp_plot<-list()

for(fs_type_i in seq_along(fs_type_opt)){
  fs_type<-fs_type_opt[fs_type_i]
  
  varimp<-readRDS(paste0("./data/model_ref/pred_in_",pred_in_d,"d_",fs_type,"_baseline_model_perf.rda"))$varimp_tbl %>%
    dplyr::mutate(suffix=gsub("_","",str_extract(Feature,"((\\_min)|(\\_slope)|(\\_change)|(\\_cum))+"))) %>%
    dplyr::mutate(feat=gsub("((\\_min)|(\\_slope)|(\\_change)|(\\_cum))+","",Feature)) %>%
    left_join(data_dict,by=c("feat"="VALUESET_ITEM")) %>%
    dplyr::mutate(feat = ifelse(is.na(VALUESET_ITEM_DESCRIPTOR),Feature,VALUESET_ITEM_DESCRIPTOR)) %>%
    dplyr::mutate(feat = ifelse(!is.na(suffix),paste0(feat,"_",suffix),feat)) %>%
    dplyr::mutate(feat_rank=paste0(ifelse(rank<10,
                                          paste0("0",rank),
                                          as.character(rank)),
                                   ".",feat)) %>%
    unique %>%
    mutate(feat_rank=paste0(substr(feat_rank,1,50),"...")) %>%
    dplyr::mutate(feat_rank=as.factor(feat_rank)) %>%
    dplyr::mutate(feat_rank=factor(feat_rank,levels=rev(levels(feat_rank)))) %>%
    mutate(pred_task=recode(pred_task,
                            `stg1up`="a.AKI>=1",
                            `stg2up`="b.AKI>=2",
                            `stg3`="c.AKI=3"))
  
  pred48_varimp_plot[[fs_type_i]]<-ggplot(varimp %>% filter(rank <= 20),
                                 aes(x=feat_rank,y=Gain_rescale))+
    geom_bar(stat="identity")+
    labs(x="Features",y="Normalized Scale",
         title=paste0("Figure",fs_type_i+10," - Top Important Variables:",fs_type))+
    coord_flip()+scale_y_continuous(trans = "reverse")+
    facet_wrap(~pred_task,scales = "free",ncol=1)
  
  final_out[[paste0("pred48_varimp_",fs_type)]]<-varimp
}
print(pred48_varimp_plot[[1]])
print(pred48_varimp_plot[[2]])

```

***

```{r final, include=F}
write.xlsx(final_out,file="./output/AKI_CDM_EXT_VALID_p2_2_Retrain_TBL.xlsx")

# ggplot sometimes create some unwanted empty .pdf file and want to clean it up
if(file.exists("./Rplots.pdf")){
  file.remove("./Rplots.pdf")
}

rm(list=ls())
gc()
```





