---
title: "Building and Validating Predictive Models for Acute Kidney Injury (AKI) using PCORnet CDM (Part II.1)"
author: "xing song"
date: "Feburary 09, 2019"
output: html_document
---
### Stage 2.1: Predictive Models Validation (Benchmark)   

In this experiment, we will validate the benchmark predictive model by quasi-replicating the model by [*Koyner et al*] for AKI risk prediction on the adult inpatients, which was trained on KUMC's data using PCORnet common data model. This learning frame work by Koyner et al is, by far, the best reported general inpatients AKI prediction model, which is based on a tuned gradient boosting machine in a discrete time survival analysis framework with pre-selected EHR data elements. We aim to extend this benchmark prediction model by including novel elements from PCORnet CDM, while effectively automate the manual efforts of feature pre-selection and cleaning. The primary outcomes of this study were the development of AKI stage ≥ 1, AKI stage ≥ 2, and AKI stage 3 in the next 24 and 48 hours. KUMC's model and some auxillary data will be delivered via securefile. 

[*Koyner et al*] https://www.ncbi.nlm.nih.gov/pubmed/29596073 

***

```{r setup, include=F}
#source utility functions
source("./R/util.R")
source("./R/var_etl_surv.R")

require_libraries(c("tidyr",
                    "dplyr",
                    "magrittr",
                    "stringr",
                    "broom",
                    "Matrix",
                    "xgboost",
                    "ROCR",
                    "PRROC",
                    "ResourceSelection",
                    "knitr",
                    "kableExtra",
                    "ggplot2",
                    "openxlsx"))

# experimental design parameters
#----prediction ending point
pred_end<-7

#-----prediction point
pred_in_d_opt<-c(1,2)

#-----prediction tasks
pred_task_lst<-c("stg1up","stg2up","stg3")

#-----feature selection type
fs_type_opt<-c("no_fs","rm_scr_bun")
rm_key<-c('2160-0','38483-4','14682-9','21232-4','35203-9','44784-7','59826-8',
          '16188-5','16189-3','59826-8','35591-7','50380-5','50381-3','35592-5',
          '44784-7','11041-1','51620-3','72271-0','11042-9','51619-5','35203-9','14682-9',
          '12966-8','12965-0','6299-2','59570-2','12964-3','49071-4','72270-2',
          '11065-0','3094-0','35234-4','14937-7',
          '3097-3','44734-2','BUN_SCR')
```


#### Preparation    

By running `Part I` of "render_report.R", the raw data tables should have already been collected and saved in the local `./data` folder (Note: these data tables are not visiable in the github ./data folder, but should be visible in the corresponding folder locally), that are

* `Table1.rda`: AKI patieht cohort table;       

* `AKI_DEMO.rda`: CDM demographic table cut for AKI cohort;   

* `AKI_VITAL.rda`: CDM vital table cut for AKI cohort;    

* `AKI_LAB.rda`: CDM lab table cut for AKI cohort;    

* `AKI_DX.rda`: CDM diagnosis table cut for AKI cohort;   

* `AKI_PX.rda`: CDM procedure table cut for AKI cohort;   

* `AKI_MED.rda`: CDM prescribing medication table cut for AKI cohort;   


***

#### Objective 2.1: Data Cleaning and Representation

In this section, the raw data tables will be cleaned and transformed to a discrete-survival-like representation, which will be used in the final modeling stage. To reduce the burden on memory requirments, the ETL (extract, transform, load) process will be performed in chunks with respect to **distinct prediction task, encounter years and variable types**. Meanwhile, indices for random paritioning will be assigned to each encounter. The ETL progress will be reported as follows:


```{r preprocess, include=F}
tbl1<-readRDS("./data//Table1.rda") %>%
  dplyr::mutate(yr=as.numeric(format(strptime(ADMIT_DATE, "%Y-%m-%d %H:%M:%S"),"%Y")))

for(pred_in_d in pred_in_d_opt){
  #--determine update time window
  tw<-as.double(seq(0,7))
  if(pred_in_d>1){
  tw<-tw[-seq_len(pred_in_d-1)]
  } 
  
  #--by chunks: encounter year
enc_yr<-tbl1 %>%
  dplyr::select(yr) %>%
  unique %>% arrange(yr) %>%
  filter(yr>2009) %>%
  unlist

#--by variable type
var_type<-c("demo","vital","lab","dx","px","med")

#--save results as array
for(pred_task in pred_task_lst){
  start_tsk<-Sys.time()
  cat("Start variable collection for task",pred_task,".\n")
  #---------------------------------------------------------------------------------------------
  
  var_by_yr<-list()
  var_bm<-list()
  rsample_idx<-c()
  for(i in seq_along(enc_yr)){
    start_i<-Sys.time()
    cat("...start variable collection for year",enc_yr[i],".\n")
    
    #--collect end_points
    dat_i<-tbl1 %>% filter(yr==enc_yr[i]) %>%
      dplyr::select(ENCOUNTERID,yr,
                    NONAKI_SINCE_ADMIT,
                    AKI1_SINCE_ADMIT,
                    AKI2_SINCE_ADMIT,
                    AKI3_SINCE_ADMIT) %>%
      gather(y,dsa_y,-ENCOUNTERID,-yr) %>%
      filter(!is.na(dsa_y)) %>%
      dplyr::mutate(y=recode(y,
                             "NONAKI_SINCE_ADMIT"=0,
                             "AKI1_SINCE_ADMIT"=1,
                             "AKI2_SINCE_ADMIT"=2,
                             "AKI3_SINCE_ADMIT"=3)) %>%
      dplyr::mutate(y=as.numeric(y))
    
    if(pred_task=="stg1up"){
      dat_i %<>%
        dplyr::mutate(y=as.numeric(y>0)) %>%
        group_by(ENCOUNTERID) %>% top_n(n=1L,wt=dsa_y) %>% ungroup
    }else if(pred_task=="stg2up"){
      dat_i %<>%
        # filter(y!=1) %>% # remove stage 1
        dplyr::mutate(y=as.numeric(y>1)) %>%
        group_by(ENCOUNTERID) %>% top_n(n=1L,wt=dsa_y) %>% ungroup
    }else if(pred_task=="stg3"){
      dat_i %<>%
        # filter(!(y %in% c(1,2))) %>% # remove stage 1,2
        dplyr::mutate(y=as.numeric(y>2)) %>%
        group_by(ENCOUNTERID) %>% top_n(n=1L,wt=dsa_y) %>% ungroup
    }else{
      stop("prediction task is not valid!")
    }
    
    #--random sampling
    rsample_idx %<>%
      bind_rows(dat_i %>% 
                  dplyr::select(ENCOUNTERID,yr) %>%
                  unique %>%
                  dplyr::mutate(cv10_idx=sample(1:10,n(),replace=T)))
    
    #--ETL variables
    X_surv<-c()
    y_surv<-c()
    var_etl_bm<-c()
    for(v in seq_along(var_type)){
      start_v<-Sys.time()
      
      #extract
      var_v<-readRDS(paste0("./data/AKI_",toupper(var_type[v]),".rda")) %>%
        semi_join(dat_i,by="ENCOUNTERID")
      
      if(var_type[v] != "demo"){
        if(var_type[v] == "med"){
          var_v %<>% 
            transform(value=strsplit(value,","),
                      dsa=strsplit(dsa,",")) %>%
            unnest(value,dsa) %>%
            dplyr::mutate(value=as.numeric(value),
                          dsa=as.numeric(dsa))
        }
        var_v %<>% filter(dsa <= pred_end)
      }
      
      #transform
      var_v<-format_data(dat=var_v,
                         type=var_type[v],
                         pred_end=pred_end)
      
      Xy_surv<-get_dsurv_temporal(dat=var_v,
                                  censor=dat_i,
                                  tw=tw,
                                  pred_in_d=pred_in_d)
      
      #load
      X_surv %<>% bind_rows(Xy_surv$X_surv) %>% unique
      y_surv %<>% bind_rows(Xy_surv$y_surv) %>% unique
      
      lapse_v<-Sys.time()-start_v
      var_etl_bm<-c(var_etl_bm,paste0(lapse_v,units(lapse_v)))
      cat("\n......finished ETL",var_type[v],"for year",enc_yr[i],"in",lapse_v,units(lapse_v),".\n")
    }
    var_by_yr[[i]]<-list(X_surv=X_surv,
                         y_surv=y_surv)
    
    lapse_i<-Sys.time()-start_i
    var_etl_bm<-c(var_etl_bm,paste0(lapse_i,units(lapse_i)))
    cat("\n...finished variabl collection for year",enc_yr[i],"in",lapse_i,units(lapse_i),".\n")
    flush.console() 
    
    var_bm[[i]]<-data.frame(bm_nm=c(var_type,"overall"),
                            bm_time=var_etl_bm,
                            stringsAsFactors = F)
  }
  #--save preprocessed data
  saveRDS(rsample_idx,file=paste0("./data/preproc/",pred_in_d,"d_rsample_idx_",pred_task,".rda"))
  saveRDS(var_by_yr,file=paste0("./data/preproc/",pred_in_d,"d_var_by_yr_",pred_task,".rda"))
  saveRDS(var_bm,file=paste0("./data/preproc/",pred_in_d,"d_var_bm",pred_task,".rda"))
  
  #---------------------------------------------------------------------------------------------
  lapse_tsk<-Sys.time()-start_tsk
  cat("\nFinish variable ETL for task:",pred_task,"in",pred_in_d,"days",",in",lapse_tsk,units(lapse_tsk),".\n")
  }
}
```

The final preprocessed intermediate tables from this code chunk should be found in the `./data/preproc/...` folder as the following intermediate data tables for different prediction tasks:   

* For AKI stage ≥ 1 in 24 hours: `1d_rsample_idx_stg1up.rda`, `1d_var_by_yr_stg1up.rda`, `1d_var_bm_stg1up.rda`   

* For AKI stage ≥ 2 in 24 hours: `1d_rsample_idx_stg2up.rda`, `1d_var_by_yr_stg2up.rda`, `1d_var_bm_stg2up.rda`   
 
* For AKI stage = 3 in 24 hours: `1d_rsample_idx_stg3.rda`, `1d_var_by_yr_stg3.rda`, `1d_var_bm_stg3.rda`   

* For AKI stage ≥ 1 in 48 hours: `2d_rsample_idx_stg1up.rda`, `2d_var_by_yr_stg1up.rda`, `1d_var_bm_stg1up.rda`   
 
* For AKI stage ≥ 2 in 48 hours: `2d_rsample_idx_stg2up.rda`, `2d_var_by_yr_stg2up.rda`, `1d_var_bm_stg2up.rda`   
 
* For AKI stage = 3 in 48 hours: `2d_rsample_idx_stg3.rda`, `2d_var_by_yr_stg3.rda`, `1d_var_bm_stg3.rda`   


***

#### Objective 2.2: Benchmark Model Validation

After receiving the `model_kumc.zip` from KUMC via securefile, unzip the data and save under `.../data/model_kumc`. The model will be validated on the overall cohort from the site by generating a AKI risk score for different AKI stages. The validation progress will be reported as follows:

```{r benchmark, echo=F}
rm(list=c("X_surv","y_surv","dat_i")); gc() #release some memory

for(pred_in_d in pred_in_d_opt){
  
  for(pred_task in pred_task_lst){
    start_tsk<-Sys.time()
    #---------------------------------------------------------------------------------------------
    var_by_task<-readRDS(paste0("./data/preproc/",pred_in_d,"d_var_by_yr_",pred_task,".rda"))
    
    for(fs_type in fs_type_opt){
      #--prepare testing set
      yr_rg<-seq(2010,2018)
      X_ts<-c()
      y_ts<-c()
      for(i in seq_along(yr_rg)){
        var_by_yr<-var_by_task[[i]]
        
        X_ts %<>% bind_rows(var_by_yr[["X_surv"]])
        y_ts %<>% bind_rows(var_by_yr[["y_surv"]])
        
        # cat(pred_in_d,",",pred_in_d,"...finish stack data of encounters from year",yr_rg[i],".\n")
      }
      
      #--pre-filter
      if(fs_type=="rm_scr_bun"){
        X_ts %<>%
          filter(!(key %in% c(rm_key,paste0(rm_key,"_change"))))
      }
      
      #--collect variables used in training
      ref_mod<-readRDS(paste0("./data/model_kumc/pred_in_",pred_in_d,"d_model_gbm_",fs_type,"_",pred_task,".rda"))
      tr_key<-data.frame(key = ref_mod$feature_names,
                         stringsAsFactors = F)
      
      #--transform testing matrix
      y_ts %<>%
        arrange(ENCOUNTERID,dsa_y) %>%
        unite("ROW_ID",c("ENCOUNTERID","dsa_y")) %>%
        arrange(ROW_ID) %>%
        unique
      
      X_ts %<>% 
        unite("ROW_ID",c("ENCOUNTERID","dsa_y")) %>%
        semi_join(y_ts,by="ROW_ID") %>%
        semi_join(tr_key,by="key")
      
      x_add<-tr_key %>%
        anti_join(data.frame(key = unique(X_ts$key),
                             stringsAsFactors = F),
                  by="key")
      
      #align with training
      if(nrow(x_add)>0){
        X_ts %<>%
          arrange(ROW_ID) %>%
          bind_rows(data.frame(ROW_ID = rep("0_0",nrow(x_add)),
                               dsa = -99,
                               key = x_add$key,
                               value = 0,
                               stringsAsFactors=F))
      }
      X_ts %<>%
        long_to_sparse_matrix(df=.,
                              id="ROW_ID",
                              variable="key",
                              val="value")
      if(nrow(x_add)>0){
        X_ts<-X_ts[-1,]
      }
      
      #check alignment
      all(row.names(X_ts)==y_ts$ROW_ID)
      all(ref_mod$feature_names==colnames(X_ts))
      
      #--covert to xgb data frame
      dtest<-xgb.DMatrix(data=X_ts,label=y_ts$y)
      
      #--validation
      valid<-data.frame(y_ts,
                        pred = predict(ref_mod,dtest),
                        stringsAsFactors = F)
      
      #--save model and other results
      saveRDS(valid,file=paste0("./data/model_kumc/pred_in_",pred_in_d,"d_valid_gbm_",fs_type,"_",pred_task,".rda"))
      #-------------------------------------------------------------------------------------------------------------
      lapse_tsk<-Sys.time()-start_tsk
      cat("\nFinish validating benchmark models for task:",pred_task,"in",pred_in_d,"with",fs_type,",in",lapse_tsk,units(lapse_tsk),".\n")
    }
  }
}

```

For each prediction task, defined as "predict AKI stage X in Y days, with/without Scr", 1 intermedicate data file have been generated and saved in `./data/model_kumc/...`, which are: 

* `..._valid_gbm_...rda`: the predictted probability on validataion set;      


***


#### Objective 2.3: Performance Evaluations for Benchmark Model

```{r perf_summ, include=F}
rm(list=c("X_ts","y_ts","dtest")); gc() #release some memory

pred_in_d_opt<-c(1,2)
fs_type_opt<-c("no_fs","rm_scr_bun")
pred_task<-c("stg1up","stg2up","stg3")

for(pred_in_d in pred_in_d_opt){
  for(fs_type in fs_type_opt){
    perf_tbl_full<-c()
    perf_tbl<-c()
    calib_tbl<-c()
    for(i in seq_along(pred_task)){
      valid<-readRDS(paste0("./data/model_kumc/pred_in_",pred_in_d,"d_valid_gbm_",fs_type,"_",pred_task[i],".rda"))
      
      #overall summary
      perf_summ<-get_perf_summ(pred=valid$pred,
                               real=valid$y,
                               keep_all_cutoffs=T)
      perf_tbl_full %<>% 
        bind_rows(perf_summ$perf_at %>% 
                    dplyr::mutate(pred_task=pred_task[i],pred_in_d=pred_in_d,fs_type=fs_type))
      
      perf_tbl %<>% 
        bind_rows(perf_summ$perf_summ %>% 
                    dplyr::mutate(pred_task=pred_task[i],pred_in_d=pred_in_d,fs_type=fs_type))
      
      #calibration
      calib<-get_calibr(pred=valid$pred,
                       real=valid$y,
                       n_bin=20)
      
      calib_tbl %<>% 
        bind_rows(calib %>% 
                    dplyr::mutate(pred_task=pred_task[i],pred_in_d=pred_in_d,fs_type=fs_type))
    }
    
    perf_out<-list(perf_tbl_full=perf_tbl_full,
                   perf_tbl=perf_tbl,
                   calib_tbl=calib_tbl)
    
    #save results as r data.frame
    saveRDS(perf_out,file=paste0("./data/model_kumc/pred_in_",pred_in_d,"d_",fs_type,"_baseline_model_perf.rda"))
  }
}
```


```{r data_dict, include=F}
tbl1<-readRDS("./data/Table1.rda") %>%
  dplyr::select(ENCOUNTERID,SERUM_CREAT_BASE) %>%
  inner_join(readRDS("./data/AKI_DEMO.rda") %>%
              filter(key=="AGE") %>%
              dplyr::select(ENCOUNTERID,key,value) %>%
              spread(key,value) %>%
              mutate(AGE=as.numeric(AGE)),
            by="ENCOUNTERID")

final_out<-list()
```


#### Task I: 24-hour advanced prediction

```{r pred24hr, include=F}
pred_in_d<-1
```


*(I.a) Overall Prediction Performance and Subgroup Analysis*   

```{r pred24_summ,echo=F}
fs_type_opt<-c("no_fs","rm_scr_bun")

#overall summary
perf_overall<-c()
for(fs_type in fs_type_opt){
  perf_overall %<>%
    bind_rows(readRDS(paste0("./data/model_kumc/pred_in_",pred_in_d,"d_",fs_type,"_baseline_model_perf.rda"))$perf_tbl %>%
                filter(overall_meas %in% c("roauc",
                                           "roauc_low",
                                           "roauc_up",
                                           "prauc1",
                                           "opt_sens",
                                           "opt_spec",
                                           "opt_ppv",
                                           "opt_npv")) %>%
                dplyr::mutate(overall_meas=recode(overall_meas,
                                                  roauc="1.ROAUC",
                                                  prauc1="2.PRAUC",
                                                  opt_sens="3.Optimal Sensitivity",
                                                  opt_spec="4.Optimal Specificity",
                                                  opt_ppv="5.Optimal Positive Predictive Value",
                                                  opt_npv="6.Optimal Negative Predictive Value"),
                              grp=paste0("Overall:",fs_type)))
}

#subgroup-age
fs_type<-"no_fs"
subgrp_age<-c()
for(i in seq_along(pred_task)){
      valid<-readRDS(paste0("./data/model_kumc/pred_in_",pred_in_d,"d_valid_gbm_",fs_type,"_",pred_task[i],".rda")) %>%
        mutate(ENCOUNTERID=gsub("_.*","",ROW_ID),
               day_at=as.numeric(gsub(".*_","",ROW_ID))) %>%
        inner_join(tbl1,by="ENCOUNTERID") %>%
        dplyr::select(ENCOUNTERID,day_at,y,pred,AGE) %>%
        dplyr::mutate(age=cut(AGE,breaks=c(0,45,65,Inf),include.lowest=T,right=F))
      
      grp_vec<-unique(valid$age)
      for(grp in seq_along(grp_vec)){
        valid_grp<-valid %>% filter(age==grp_vec[grp])
        grp_summ<-get_perf_summ(pred=valid_grp$pred,
                                real=valid_grp$y,
                                keep_all_cutoffs=F)$perf_summ %>%
          dplyr::filter(overall_meas %in% c("roauc",
                                            "roauc_low",
                                            "roauc_up",
                                            "prauc1",
                                            "opt_sens",
                                            "opt_spec",
                                            "opt_ppv",
                                            "opt_npv")) %>%
          dplyr::mutate(overall_meas=recode(overall_meas,
                                            roauc="1.ROAUC",
                                            prauc1="2.PRAUC",
                                            opt_sens="3.Optimal Sensitivity",
                                            opt_spec="4.Optimal Specificity",
                                            opt_ppv="5.Optimal Positive Predictive Value",
                                            opt_npv="6.Optimal Negative Predictive Value"))
        subgrp_age %<>% 
          bind_rows(grp_summ %>% 
                      dplyr::mutate(grp=paste0("Subgrp_AGE:",grp_vec[grp]),
                                    pred_task=paste0("stg",i,"up")))
      }
  }
  
  #subgroup-scr
  subgrp_scr<-c()
  for(i in seq_along(pred_task)){
      valid<-readRDS(paste0("./data/model_kumc/pred_in_",pred_in_d,"d_valid_gbm_",fs_type,"_",pred_task[i],".rda")) %>%
        mutate(ENCOUNTERID=gsub("_.*","",ROW_ID),
               day_at=as.numeric(gsub(".*_","",ROW_ID))) %>%
        inner_join(tbl1,by="ENCOUNTERID") %>%
        dplyr::select(ENCOUNTERID,day_at,y,pred,SERUM_CREAT_BASE) %>%
        mutate(admit_scr=cut(SERUM_CREAT_BASE,breaks=c(0,1,2,3,Inf),include.lowest=T,right=F))
      
      grp_vec<-unique(valid$admit_scr)
      for(grp in seq_along(grp_vec)){
        valid_grp<-valid %>% filter(admit_scr==grp_vec[grp])
        grp_summ<-get_perf_summ(pred=valid_grp$pred,
                                real=valid_grp$y,
                                keep_all_cutoffs=F)$perf_summ %>%
          filter(overall_meas %in% c("roauc",
                                     "roauc_low",
                                     "roauc_up",
                                     "prauc1",
                                     "opt_sens",
                                     "opt_spec",
                                     "opt_ppv",
                                     "opt_npv")) %>%
          dplyr::mutate(overall_meas=recode(overall_meas,
                                            roauc="1.ROAUC",
                                            prauc1="2.PRAUC",
                                            opt_sens="3.Optimal Sensitivity",
                                            opt_spec="4.Optimal Specificity",
                                            opt_ppv="5.Optimal Positive Predictive Value",
                                            opt_npv="6.Optimal Negative Predictive Value"))
        subgrp_scr %<>% 
          bind_rows(grp_summ %>% 
                      dplyr::mutate(grp=paste0("Subgrp_Scr_Base:",grp_vec[grp]),
                                    pred_task=paste0("stg",i,"up")))
      }
  }

perf_overall %<>%
  bind_rows(subgrp_age %>% 
              dplyr::mutate(pred_in_d=pred_in_d,fs_type=fs_type)) %>%
  bind_rows(subgrp_scr %>% 
              dplyr::mutate(pred_in_d=pred_in_d,fs_type=fs_type)) %>%
  dplyr::select(pred_in_d,fs_type,grp,pred_task,overall_meas,meas_val) %>%
  dplyr::mutate(pred_task=recode(pred_task,
                                 `stg1up`="a.AKI>=1",
                                 `stg2up`="b.AKI>=2",
                                 `stg3`="c.AKI=3",
                                 `stg3up`="c.AKI=3"),
                meas_val=round(meas_val,4)) %>%
                spread(overall_meas,meas_val) %>%
                dplyr::mutate(roauc_up=paste0(roauc_up,")")) %>%
                unite("1.ROAUC_CI",c("roauc_low","roauc_up"),sep=",") %>%
                unite("1.ROAUC",c("1.ROAUC","1.ROAUC_CI"),sep=" (")  

row_grp_pos<-perf_overall %>% 
  arrange(grp,pred_task) %>%
  mutate(rn=1:n()) %>%
  dplyr::mutate(root_grp=gsub(":.*","",grp)) %>%
  group_by(root_grp,pred_in_d) %>%
  dplyr::summarize(begin=rn[1],
                   end=rn[n()]) %>%
  ungroup

kable(perf_overall %>% arrange(grp,pred_task),
      caption="Table1 - 24-Hour Prediction of AKI1, AKI2, AKI3") %>%
  kable_styling("striped", full_width = F) %>%
  group_rows("Overall", row_grp_pos$begin[1],row_grp_pos$end[1]) %>%
  group_rows("Subgroup-Age", row_grp_pos$begin[2],row_grp_pos$end[2]) %>%
  group_rows("Subgroup-Scr_Base", row_grp_pos$begin[3],row_grp_pos$end[3])

final_out[["pred24_summ"]]<-perf_overall
```


*(I.b) Full metric paths at various cutoff points*

```{r pred24_cutoff, echo=F,fig.height = 5, fig.width = 12}
#plot out sens, spec, ppv, npv on a scale of cutoff probabilities
brks<-unique(c(0,seq(0.001,0.01,by=0.001),seq(0.02,0.1,by=0.01),seq(0.2,1,by=0.1)))
pred24_cutoff_plot<-list()

for(fs_type_i in seq_along(fs_type_opt)){
  fs_type<-fs_type_opt[fs_type_i]
  
  perf_cutoff<-readRDS(paste0("./data/model_kumc/pred_in_",pred_in_d,"d_",fs_type,"_baseline_model_perf.rda"))$perf_tbl_full %>%
    dplyr::select(cutoff,rec_sens,spec,ppv,npv,pred_task) %>%
    mutate(bin=cut(cutoff,breaks=brks,include.lowest=T,label=F)) %>%
    group_by(bin,pred_task) %>%
    dplyr::summarise(cutoff=round(max(cutoff),3),
                     sens_l=quantile(rec_sens,0.025,na.rm=T),
                     sens=median(rec_sens,na.rm=T),
                     sens_u=quantile(rec_sens,0.975,na.rm=T),
                     spec_l=quantile(spec,0.025,na.rm=T),
                     spec=median(spec,na.rm=T),
                     spec_u=quantile(spec,0.975,na.rm=T),
                     ppv_l=quantile(ppv,0.025,na.rm=T),
                     ppv=median(ppv,na.rm=T),
                     ppv_u=quantile(ppv,0.975,na.rm=T),
                     npv_l=quantile(npv,0.025,na.rm=T),
                     npv=median(npv,na.rm=T),
                     npv_u=quantile(npv,0.975,na.rm=T)) %>%
    ungroup %>% dplyr::select(-bin) %>%
    gather(metric_type,metric_val,-cutoff,-pred_task) %>%
    mutate(metric_type2=ifelse(grepl("_l",metric_type),"low",
                               ifelse(grepl("_u",metric_type),"up","mid")),
          metric_type=gsub("_.*","",metric_type)) %>%
    spread(metric_type2,metric_val) %>%
    mutate(pred_task=recode(pred_task,
                            `stg1up`="a.AKI>=1",
                            `stg2up`="b.AKI>=2",
                            `stg3`="c.AKI=3"))
  
  pred24_cutoff_plot[[fs_type_i]]<-ggplot(perf_cutoff %>% dplyr::filter(cutoff <=0.15),
                   aes(x=cutoff,y=mid,color=metric_type,fill=metric_type))+
    geom_line()+ geom_ribbon(aes(ymin=low,ymax=up),alpha=0.3)+
    labs(x="cutoff probability",y="performance metrics",
         title=paste0("Figure",fs_type_i," - Metrics at different cutoff points:",fs_type))+
    facet_wrap(~pred_task,scales="free",ncol=3)
  
  final_out[[paste0("pred24_cutoff_",fs_type)]]<-perf_cutoff
}

print(pred24_cutoff_plot[[1]])
print(pred24_cutoff_plot[[2]])
```


*(I.c) Calibration Plots*

```{r pred24_calib, echo=F, fig.height = 5, fig.width = 12}
pred24_calib_plot<-list()

# plot calibration
for(fs_type_i in seq_along(fs_type_opt)){
  fs_type<-fs_type_opt[fs_type_i]
  
  calib_tbl<-readRDS(paste0("./data/model_kumc/pred_in_",pred_in_d,"d_",fs_type,"_baseline_model_perf.rda"))$calib_tbl %>%
    mutate(pred_task=recode(pred_task,
                          `stg1up`="a.AKI>=1",
                          `stg2up`="b.AKI>=2",
                          `stg3`="c.AKI=3"))
  
  pred24_calib_plot[[fs_type_i]]<-ggplot(calib_tbl,aes(x=y_p,y=pred_p))+
    geom_point()+geom_abline(intercept=0,slope=1)+
    geom_errorbar(aes(ymin=binCI_lower,ymax=binCI_upper))+
    labs(x="Actual Probability",y="Predicted Probability",
         title=paste0("Figure",fs_type_i+2," - Calibration:",fs_type))+
    facet_wrap(~pred_task,scales="free")
  
  final_out[[paste0("pred24_calibr_",fs_type)]]<-calib_tbl
}
print(pred24_calib_plot[[1]])
print(pred24_calib_plot[[2]])
```


***


#### Task II: 48-hour advanced prediction

```{r pred48hr, include=F}
pred_in_d<-2
```

*(II.a) Overall Prediction Performance*

```{r pred48_summ,echo=F}
fs_type_opt<-c("no_fs","rm_scr_bun")

#overall summary
perf_overall<-c()
for(fs_type in fs_type_opt){
  perf_overall %<>%
    bind_rows(readRDS(paste0("./data/model_kumc/pred_in_",pred_in_d,"d_",fs_type,"_baseline_model_perf.rda"))$perf_tbl %>%
                filter(overall_meas %in% c("roauc",
                                           "roauc_low",
                                           "roauc_up",
                                           "prauc1",
                                           "opt_sens",
                                           "opt_spec",
                                           "opt_ppv",
                                           "opt_npv")) %>%
                dplyr::mutate(overall_meas=recode(overall_meas,
                                                  roauc="1.ROAUC",
                                                  prauc1="2.PRAUC",
                                                  opt_sens="3.Optimal Sensitivity",
                                                  opt_spec="4.Optimal Specificity",
                                                  opt_ppv="5.Optimal Positive Predictive Value",
                                                  opt_npv="6.Optimal Negative Predictive Value"),
                              grp=paste0("Overall:",fs_type)))
}

#subgroup-age
fst_type<-"no_fs"
subgrp_age<-c()
for(i in seq_along(pred_task)){
      valid<-readRDS(paste0("./data/model_kumc/pred_in_",pred_in_d,"d_valid_gbm_",fs_type,"_",pred_task[i],".rda")) %>%
        mutate(ENCOUNTERID=gsub("_.*","",ROW_ID),
               day_at=as.numeric(gsub(".*_","",ROW_ID))) %>%
        inner_join(tbl1,by="ENCOUNTERID") %>%
        dplyr::select(ENCOUNTERID,day_at,y,pred,AGE) %>%
        dplyr::mutate(age=cut(AGE,breaks=c(0,45,65,Inf),include.lowest=T,right=F))
      
      grp_vec<-unique(valid$age)
      for(grp in seq_along(grp_vec)){
        valid_grp<-valid %>% filter(age==grp_vec[grp])
        grp_summ<-get_perf_summ(pred=valid_grp$pred,
                                real=valid_grp$y,
                                keep_all_cutoffs=F)$perf_summ %>%
          dplyr::filter(overall_meas %in% c("roauc",
                                            "roauc_low",
                                            "roauc_up",
                                            "prauc1",
                                            "opt_sens",
                                            "opt_spec",
                                            "opt_ppv",
                                            "opt_npv")) %>%
          dplyr::mutate(overall_meas=recode(overall_meas,
                                            roauc="1.ROAUC",
                                            prauc1="2.PRAUC",
                                            opt_sens="3.Optimal Sensitivity",
                                            opt_spec="4.Optimal Specificity",
                                            opt_ppv="5.Optimal Positive Predictive Value",
                                            opt_npv="6.Optimal Negative Predictive Value"))
        subgrp_age %<>% 
          bind_rows(grp_summ %>% 
                      dplyr::mutate(grp=paste0("Subgrp_AGE:",grp_vec[grp]),
                                    pred_task=paste0("stg",i,"up")))
      }
  }
  
  #subgroup-scr
  subgrp_scr<-c()
  for(i in seq_along(pred_task)){
      valid<-readRDS(paste0("./data/model_kumc/pred_in_",pred_in_d,"d_valid_gbm_",fs_type,"_",pred_task[i],".rda")) %>%
        mutate(ENCOUNTERID=gsub("_.*","",ROW_ID),
               day_at=as.numeric(gsub(".*_","",ROW_ID))) %>%
        inner_join(tbl1,by="ENCOUNTERID") %>%
        dplyr::select(ENCOUNTERID,day_at,y,pred,SERUM_CREAT_BASE) %>%
        mutate(admit_scr=cut(SERUM_CREAT_BASE,breaks=c(0,1,2,3,Inf),include.lowest=T,right=F))
      
      grp_vec<-unique(valid$admit_scr)
      for(grp in seq_along(grp_vec)){
        valid_grp<-valid %>% filter(admit_scr==grp_vec[grp])
        grp_summ<-get_perf_summ(pred=valid_grp$pred,
                                real=valid_grp$y,
                                keep_all_cutoffs=F)$perf_summ %>%
          filter(overall_meas %in% c("roauc",
                                     "roauc_low",
                                     "roauc_up",
                                     "prauc1",
                                     "opt_sens",
                                     "opt_spec",
                                     "opt_ppv",
                                     "opt_npv")) %>%
          dplyr::mutate(overall_meas=recode(overall_meas,
                                            roauc="1.ROAUC",
                                            prauc1="2.PRAUC",
                                            opt_sens="3.Optimal Sensitivity",
                                            opt_spec="4.Optimal Specificity",
                                            opt_ppv="5.Optimal Positive Predictive Value",
                                            opt_npv="6.Optimal Negative Predictive Value"))
        subgrp_scr %<>% 
          bind_rows(grp_summ %>% 
                      dplyr::mutate(grp=paste0("Subgrp_Scr_Base:",grp_vec[grp]),
                                    pred_task=paste0("stg",i,"up")))
      }
  }

perf_overall %<>%
  bind_rows(subgrp_age %>% 
              dplyr::mutate(pred_in_d=pred_in_d,fs_type=fs_type)) %>%
  bind_rows(subgrp_scr %>% 
              dplyr::mutate(pred_in_d=pred_in_d,fs_type=fs_type)) %>%
  dplyr::select(pred_in_d,fs_type,grp,pred_task,overall_meas,meas_val) %>%
  dplyr::mutate(pred_task=recode(pred_task,
                                 `stg1up`="a.AKI>=1",
                                 `stg2up`="b.AKI>=2",
                                 `stg3`="c.AKI=3",
                                 `stg3up`="c.AKI=3"),
                meas_val=round(meas_val,4)) %>%
                spread(overall_meas,meas_val) %>%
                dplyr::mutate(roauc_up=paste0(roauc_up,")")) %>%
                unite("1.ROAUC_CI",c("roauc_low","roauc_up"),sep=",") %>%
                unite("1.ROAUC",c("1.ROAUC","1.ROAUC_CI"),sep=" (")  

row_grp_pos<-perf_overall %>% 
  arrange(grp,pred_task) %>%
  mutate(rn=1:n()) %>%
  dplyr::mutate(root_grp=gsub(":.*","",grp)) %>%
  group_by(root_grp,pred_in_d) %>%
  dplyr::summarize(begin=rn[1],
                   end=rn[n()]) %>%
  ungroup

kable(perf_overall %>% arrange(grp,pred_task),
      caption="Table1 - 48-Hour Prediction of AKI1, AKI2, AKI3") %>%
  kable_styling("striped", full_width = F) %>%
  group_rows("Overall", row_grp_pos$begin[1],row_grp_pos$end[1]) %>%
  group_rows("Subgroup-Age", row_grp_pos$begin[2],row_grp_pos$end[2]) %>%
  group_rows("Subgroup-Scr_Base", row_grp_pos$begin[3],row_grp_pos$end[3])

final_out[["pred48_summ"]]<-perf_overall
```


*(II.b) Full metric paths at various cutoff points*

```{r pred48_cutoff,echo=F, fig.height = 5, fig.width = 12}
pred48_cutoff_plot<-list()

#plot out sens, spec, ppv, npv on a scale of cutoff probabilities
brks<-unique(c(0,seq(0.001,0.01,by=0.001),seq(0.02,0.1,by=0.01),seq(0.2,1,by=0.1)))
for(fs_type_i in seq_along(fs_type_opt)){
  fs_type<-fs_type_opt[fs_type_i]
  
  perf_cutoff<-readRDS(paste0("./data/model_kumc/pred_in_",pred_in_d,"d_",fs_type,"_baseline_model_perf.rda"))$perf_tbl_full %>%
  dplyr::select(cutoff,size,rec_sens,spec,ppv,npv,pred_task) %>%
  mutate(bin=cut(cutoff,breaks=brks,include.lowest=T,label=F)) %>%
  group_by(bin,pred_task) %>%
  dplyr::summarise(size=sum(size),
                   cutoff=round(max(cutoff),3),
                   sens_l=quantile(rec_sens,0.025,na.rm=T),
                   sens=median(rec_sens,na.rm=T),
                   sens_u=quantile(rec_sens,0.975,na.rm=T),
                   spec_l=quantile(spec,0.025,na.rm=T),
                   spec=median(spec,na.rm=T),
                   spec_u=quantile(spec,0.975,na.rm=T),
                   ppv_l=quantile(ppv,0.025,na.rm=T),
                   ppv=median(ppv,na.rm=T),
                   ppv_u=quantile(ppv,0.975,na.rm=T),
                   npv_l=quantile(npv,0.025,na.rm=T),
                   npv=median(npv,na.rm=T),
                   npv_u=quantile(npv,0.975,na.rm=T)) %>%
  ungroup %>%
  group_by(pred_task) %>%
  dplyr::mutate(size_overall=sum(size),
                size_cum=cumsum(size)) %>%
  ungroup %>%
  mutate(size_abv=size_overall-size_cum) %>%
  dplyr::select(-bin,-size_overall,-size_cum,-size) %>%
  gather(metric_type,metric_val,-cutoff,-size_abv,-pred_task) %>%
  mutate(metric_type2=ifelse(grepl("_l",metric_type),"low",
                             ifelse(grepl("_u",metric_type),"up","mid")),
         metric_type=gsub("_.*","",metric_type)) %>%
  spread(metric_type2,metric_val) %>%
  mutate(pred_task=recode(pred_task,
                          `stg1up`="a.At least AKI1",
                          `stg2up`="b.At least AKI2",
                          `stg3`="c.AKI3"))

  pred48_cutoff_plot[[fs_type_i]]<-ggplot(perf_cutoff %>% dplyr::filter(cutoff <=0.15),
                                          aes(x=cutoff,y=mid,color=metric_type,fill=metric_type))+
    geom_line()+ geom_ribbon(aes(ymin=low,ymax=up),alpha=0.3)+
    labs(x="cutoff probability",y="performance metrics",
         title=paste0("Figure",fs_type_i+6," - Metrics at different cutoff points:",fs_type))+
    facet_wrap(~pred_task,scales="free",ncol=3)
  
  final_out[[paste0("pred48_cutoff_",fs_type)]]<-perf_cutoff
}
print(pred48_cutoff_plot[[1]])
print(pred48_cutoff_plot[[2]])
```


*(II.c) Calibration Plots*

```{r pred48_calib,echo=F, fig.height = 5, fig.width = 12}
pred48_calib_plot<-list()

# plot calibration
for(fs_type_i in seq_along(fs_type_opt)){
  fs_type<-fs_type_opt[fs_type_i]
  
  calib_tbl<-readRDS(paste0("./data/model_kumc/pred_in_",pred_in_d,"d_",fs_type,"_baseline_model_perf.rda"))$calib_tbl %>%
  mutate(pred_task=recode(pred_task,
                          `stg1up`="a.AKI>=1",
                          `stg2up`="b.AKI>=2",
                          `stg3`="c.AKI=3"))
  
  pred48_calib_plot[[fs_type_i]]<-ggplot(calib_tbl,aes(x=y_p,y=pred_p))+
    geom_point()+geom_abline(intercept=0,slope=1)+
    geom_errorbar(aes(ymin=binCI_lower,ymax=binCI_upper))+
    labs(x="Actual Probability",y="Predicted Probability",
         title=paste0("Figure",fs_type_i+8," - Calibration:",fs_type))+
    facet_wrap(~pred_task,scales="free")
  
  final_out[[paste0("pred48_calibr_",fs_type)]]<-calib_tbl
}
print(pred48_calib_plot[[1]])
print(pred48_calib_plot[[2]])
```


***

```{r final, include=F}
write.xlsx(final_out,file="./output/AKI_CDM_EXT_VALID_p2_1_Benchmark_TBL.xlsx")

# ggplot sometimes create some unwanted empty .pdf file and want to clean it up
if(file.exists("./Rplots.pdf")){
  file.remove("./Rplots.pdf")
}

rm(list=ls())
gc()
```





