---
title: "AKI_CDM_EXT_VALID_p1_QA"
author: "xing song"
date: "August 23, 2018"
output: html_document
---

## Building and Validating Predictive Models for Acute Kidney Injury (AKI) using PCORnet CDM ##

### Aim 1: Data Feasibility Evaluation over GPC sites. ###
#### Objective 1.1: Study cohort extraction and characterization ####

Eligible AKI patients are defined as:
...


AKI Stages are defined based on KDIGO:

-- AKI 1: increase in serum creatinine by >=0.3mg/dL within 48 hours OR 1.5-1.9 times baseline within 7 days

-- AKI 2: 2.0 to 2.9 times baseline within 7 days

-- AKI 3: increase in serum creatinine to >= 4.0mg/dL within 48 hours OR 3.0 times baseline within 7 days


```{r eval=F}
#establish the connection between r-studio and CDM server
config_file_path<-"../config.csv"
config_file<-read.csv(config_file_path)
conn<-dbConnect(Oracle(),
                config_file$username,
                config_file$password,
                config_file$access)

cohort<-extract_cohort(conn,
                       cdm_db_schema="PCORNET_CDM_C4R3",
                       oracle_temp_schema=cdm_db_schema,
                       start_date="2010-01-01",
                       end_date="current date")

#save Table1
Table1<-cohort$aki_enc
save(Table1,file="./data/Table1.Rdata")

#print out attrition table
print(cohort$attrition)
```

#### Objective 1.2: Variable Collection ####
```{r eval=F}

```


#### Objective 1.3: Quality Check ####
```{r eval=F}

```




