---
title: "Building and Validating Predictive Models for Acute Kidney Injury (AKI) using PCORnet CDM (Part I)"
author: "xing song"
date: "August 23, 2018"
output: html_document
---
### Aim 1: Data Feasibility Evaluation over GPC sites.
#### Objective 1.1: Study cohort extraction and characterization

##### Inclusion criteria:

* (IP,IS,EI) visits with length of stay (LOS) >= 2 days; 
* age at visit >= 18 years old

***

##### Exclusions:

* had no documented SCr measurements during admission; 
* had initial SCr greater than or equal to 1.3 mg/dL; 
* developed stage 2 AKI (eGFR <= 15 mL/min per 1.73m^2) initially; 
* pre-existing chronic renal failure (DX); 
* received RRT within 48 hours since admission; 
* burn patients (DRG)

***

##### AKI Staging

AKI Stages are defined based on [KDIGO](http://www.kdigo.org/clinical_practice_guidelines/pdf/KDIGO%20AKI%20Guideline.pdf):

* AKI 1: increase in serum creatinine by >=**0.3 mg/dL** within **48 hours** OR **1.5-1.9 times** baseline within **7 days**; 
* AKI 2: **2.0 to 2.9 times** baseline within **7 days**; 
* AKI 3: increase in serum creatinine to >= **4.0 mg/dL** within **48 hours** OR **3.0 times** baseline within **7 days**


```{r eval=F}
#establish the connection between r-studio and CDM server
config_file_path<-"../config.csv"
config_file<-read.csv(config_file_path)
conn<-dbConnect(Oracle(),
                config_file$username,
                config_file$password,
                config_file$access)

#extract cohort --Table1
cohort<-extract_cohort(conn,
                       cdm_db_schema,
                       oracle_temp_schema=cdm_db_schema,
                       start_date="2010-01-01",
                       end_date="current date")

#save Table1
Table1<-cohort$aki_enc
save(Table1,file="./data/Table1.Rdata")

#print out attrition table
print(cohort$attrition)
```

#### Objective 1.2: Variable Collection

```{r eval=F}
#collect variables
vari<-collect_variable(conn,
                       cdm_db_schema,
                       oracle_temp_schema=cdm_db_schema)

#save tables
save(vari,file="./data/vari_tbls.Rdata")
```


#### Objective 1.3: Quality Check


##### Demographics




##### Vital: HT, WT, BMI
```{r eval=F}
#vital
#detect outliers - HT, WT, BMI
ht_wt_bmi<-AKI_VITAL %>% 
  dplyr::select(PATID, VITALID, MEASURE_DATE_TIME,
                HT,WT,ORIGINAL_BMI) %>%
  dplyr::filter(!is.na(HT) & !is.na(WT)) %>%
  left_join(AKI_DEMO %>% dplyr::select(PATID,AGE,SEX) %>% unique,
            by="PATID") %>%
  dplyr::mutate(age_grp= case_when(AGE < 30 ~ 1,
                                   AGE >= 30 & AGE < 40 ~ 2,
                                   AGE >= 40 & AGE < 50 ~ 3,
                                   AGE >= 50 & AGE < 60 ~ 4,
                                   AGE >= 60 & AGE < 70 ~ 5,
                                   AGE >= 70 ~ 6)) %>%
  dplyr::select(-AGE) %>%
  gather(key,val,-PATID,-age_grp,-SEX,-VITALID,-MEASURE_DATE_TIME) %>%
  group_by(age_grp,SEX,key) %>%
  dplyr::mutate(low_bd=ifelse(key=="ORIGINAL_BMI",
                              pmax(10,median(val,na.rm=T)-3*IQR(val,na.rm=T,type=8)),
                              pmax(50,median(val,na.rm=T)-3*IQR(val,na.rm=T,type=8))),   
                up_bd=pmin(500,median(val,na.rm=T)+4.5*IQR(val,na.rm=T,type=8))) %>% 
  ungroup %>%
  dplyr::mutate(outlier_ind = ifelse(val > up_bd | val < low_bd, 1,0))

#outliers summaries
ht_wt_bmi %>% filter(outlier_ind==1) %>%
  group_by(age_grp,SEX,key,low_bd,up_bd) %>%
  dplyr::summarize(pat_cnt=length(unique(PATID))) %>% 
  ungroup %>% unique %>% arrange(age_grp,SEX,key) %>%
  View


#detect outliers - SYSTOLIC, DIASTOLIC, at encounter level
# AKI_VITAL_test<-AKI_VITAL %>% dplyr::filter(PATID < 100)
# scalability issue!
bp<-AKI_VITAL %>% 
  dplyr::select(ENCOUNTERID,VITALID, MEASURE_DATE_TIME,
                SYSTOLIC, DIASTOLIC) %>%
  dplyr::filter(!is.na(SYSTOLIC) & !is.na(DIASTOLIC)) %>%
  gather(key,val,-ENCOUNTERID,-VITALID, -MEASURE_DATE_TIME) %>%
  group_by(key,ENCOUNTERID) %>%
  dplyr::mutate(measure_time=rank(MEASURE_DATE_TIME),
                freq=n()) %>%
  dplyr::mutate(loess_fit=predict(loess(val~measure_time),se=T)$fit,
                loess_se=predict(loess(val~measure_time),se=T)$se) %>%
  dplyr::mutate(low_bd=loess_fit-1.5*sqrt(freq)*loess_se,
                up_bd=loess_fit+1.5*sqrt(freq)*loess_se) %>%
  ungroup %>%
  dplyr::mutate(outlier_ind = ifelse(val > up_bd | val < low_bd, 1,0))


#outliers summaries
bp %>% filter(outlier_ind==1) %>%
  group_by(key) %>%
  dplyr::summarize(enc_cnt=length(unique(ENCOUNTERID)),
                   low_bd=median(low_bd),
                   low_bd_max=max(low_bd),
                   up_bd=median(up_bd),
                   up_bd=min(up_bd)) %>%
  View  
```

##### Vital: Blood Pressure



##### Vital: Smoking Status




##### Labs




##### Medications





##### Diagnosis, Procedure




